---
title: "{{< iconify ph chart-line >}} Example Dashboard"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

Here, you can see how to add text within a dashboard.

## Add a new heading like this

A line break is displayed when you add a new section.

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

# Load data
data <- readRDS('dataset_2867obs.rds')

# Data summary
cat('Dataset loaded:', nrow(data), 'rows,', ncol(data), 'columns\n')

```

## Example 1: Stacked Bars


::: {.panel-tabset}

### {{< iconify ph chart-bar >}} Change the title here...


If you want to add text within the tab, you can do so here.

```{r demographics}
# Change the title here...
result <- viz_stackedbar(
  data = data,
  title = "Change the title here...",
  x_var = "degree_1a",
  stack_var = "happy_1a",
  subtitle = "You can add a subtitle using the subtitle argument",
  x_label = "Education Level",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 500)
}
result <- htmltools::div(
  style = 'height: 500px !important; min-height: 500px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "degree_1a",
        stack_var = "happy_1a",
        title = "Change the title here...",
        subtitle = "You can add a subtitle using the subtitle argument",
        x_label = "Education Level",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "If you want to add text within the tab, you can do so here.",
        text_position = "above",
        icon = "ph:chart-bar",
        height = 500,
        tabgroup = "demographics")
```
[Full documentation](../articles/tutorial_dashboard_code.html#stacked-bar-happiness-education)

</details>

## Example 1: Stacked Bars


::: {.panel-tabset}

### {{< iconify ph gender-intersex >}} Tabset #2

```{r demographics-2}
# Tabset #2
result <- viz_stackedbar(
  data = data,
  title = "Tabset #2",
  x_var = "sex_1a",
  stack_var = "happy_1a",
  subtitle = "Another example subtitle here!",
  x_label = "Gender",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 450)
}
result <- htmltools::div(
  style = 'height: 450px !important; min-height: 450px !important; width: 100%; overflow: visible;',
  result
)

result
```

Change the position of the text using the `text_position` argument.



:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "sex_1a",
        stack_var = "happy_1a",
        title = "Tabset #2",
        subtitle = "Another example subtitle here!",
        x_label = "Gender",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "Change the position of the text using the `text_position` argument.",
        text_position = "below",
        icon = "ph:gender-intersex",
        height = 450,
        tabgroup = "demographics")
```
[Full documentation](../articles/tutorial_dashboard_code.html#stacked-bar-happiness-gender)

</details>

## Example 2: Heatmap


::: {.panel-tabset}

### {{< iconify ph heatmap >}} Trust by Education and Age


Here's another example of the kind of plots you can generate in your dashboard.

```{r social}
# Trust by Education and Age
result <- viz_heatmap(
  data = data,
  title = "Trust by Education and Age",
  x_var = "degree_1a",
  y_var = "age_1a",
  value_var = "trust_1a",
  subtitle = "Average trust levels across education and age groups",
  x_label = "Example x axis",
  y_label = "Customizable y label",
  value_label = "You can change the label here too...",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 600)
}
result <- htmltools::div(
  style = 'height: 600px !important; min-height: 600px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "heatmap",
        x_var = "degree_1a",
        y_var = "age_1a",
        value_var = "trust_1a",
        title = "Trust by Education and Age",
        subtitle = "Average trust levels across education and age groups",
        x_label = "Example x axis",
        y_label = "Customizable y label",
        value_label = "You can change the label here too...",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "Heres another example of the kind of plots you can generate in your dashboard.",
        text_position = "above",
        icon = "ph:heatmap",
        height = 600,
        tabgroup = "social")
```
[Full documentation](../articles/tutorial_dashboard_code.html#heatmap-trust-education-age)

</details>

## Example 2: Heatmap


::: {.panel-tabset}

### {{< iconify ph chart-pie >}} Trust by Region and Education


Educational and regional patterns in trust distribution.

```{r social-2}
# Trust by Region and Education
result <- viz_heatmap(
  data = data,
  title = "Trust by Region and Education",
  x_var = "region_1a",
  y_var = "degree_1a",
  value_var = "trust_1a",
  subtitle = "Educational and regional patterns in trust levels",
  x_label = "Region",
  y_label = "Education Level",
  value_label = "Trust Level",
  y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 550)
}
result <- htmltools::div(
  style = 'height: 550px !important; min-height: 550px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "heatmap",
        x_var = "region_1a",
        y_var = "degree_1a",
        value_var = "trust_1a",
        title = "Trust by Region and Education",
        subtitle = "Educational and regional patterns in trust levels",
        x_label = "Region",
        y_label = "Education Level",
        value_label = "Trust Level",
        y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "Educational and regional patterns in trust distribution.",
        text_position = "above",
        icon = "ph:chart-pie",
        height = 550,
        tabgroup = "social")
```
[Full documentation](../articles/tutorial_dashboard_code.html#heatmap-trust-region-education)

</details>

