---
title: "{{< iconify ph house >}} Welcome to the Tutorial Dashboard!"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

Welcome to the **Tutorial Dashboard**, providing insights into the General Social Survey (GSS) data!

This dashboard demonstrates how to use the **dashboardr** package to create beautiful, interactive dashboards using real survey data.

The tutorial is divided into main sections:

[{{< iconify ph chart-line >}} Example Dashboard](example_dashboard.html) - Tabbed visualizations with stacked bars and heatmaps

[{{< iconify ph chart-pie >}} Standalone Charts](standalone_charts.html) - Charts without tabsets

[{{< iconify ph chalkboard-simple >}} Text Page](text_only_page.html) - Text-only content example

[{{< iconify ph link >}} Showcase Dashboard](showcase_dashboard.html) - Full feature demonstration

{{< iconify ph cursor-click-fill >}} Click on the hyperlinks above for quick navigation.

---

## Dashboard File Structure

When you create a dashboard with dashboardr, the following files are generated:

```
tutorial_dashboard/
├── _quarto.yml
├── index.qmd
├── example_dashboard.qmd
├── standalone_charts.qmd
├── text_only_page.qmd
└── showcase_dashboard.qmd
```

---

## About the Data

This dashboard uses data from the **General Social Survey (GSS)**, a nationally representative survey of adults in the United States conducted since 1972. We explore patterns in:

- **Happiness** - Self-reported happiness levels
- **Trust** - Interpersonal trust measures
- **Political attitudes** - Party identification and ideological views

## Getting Started with dashboardr

Each visualization includes a collapsible **View R Code** section showing exactly how it was created. Click the accordion to see the code and link to full documentation.

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


<details>
<summary>{{< iconify ph code >}} View Full Dashboard Code</summary>

```r
library(dashboardr)
library(dplyr)

# Load GSS data
data(gss_panel20, package = "gssr")
gss_clean <- gss_panel20 %>%
  select(age_1a, sex_1a, degree_1a, region_1a,
         happy_1a, trust_1a, fair_1a, helpful_1a,
         polviews_1a, partyid_1a, class_1a) %>%
  filter(if_any(everything(), ~ !is.na(.)))

# Create visualizations
analysis_vizzes <- create_content() %>%
  add_viz(type = "stackedbar",
          x_var = "degree_1a", stack_var = "happy_1a",
          title = "Happiness by Education",
          tabgroup = "demographics") %>%
  add_viz(type = "heatmap",
          x_var = "degree_1a", y_var = "age_1a", value_var = "trust_1a",
          title = "Trust by Education and Age",
          tabgroup = "social")

# Create dashboard
create_dashboard(
  output_dir = "my_dashboard",
  title = "My Dashboard",
  logo = "logo.png",
  theme = "flatly"
) %>%
  add_page(name = "Welcome", text = "Landing page content",
           icon = "ph:house", is_landing_page = TRUE) %>%
  add_page(name = "Analysis", data = gss_clean,
           content = analysis_vizzes, icon = "ph:chart-line") %>%
  generate_dashboard()
```
[View full tutorial_dashboard.R source code](https://github.com/favstats/dashboardr/blob/main/R/tutorial_dashboard.R)

</details>

