---
title: "{{< iconify ph chalkboard-simple-bold >}} Text & Content"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


<details>
<summary>{{< iconify ph code >}} View Full Page Code</summary>

```r
# === TEXT & CONTENT PAGE ===
# This page demonstrates markdown formatting and content blocks

# Create content with text and accordions
text_content <- create_content() %>%
  add_text(md_text(
    "This page demonstrates text formatting and content blocks.",
    "",
    "You can use **bold text** for emphasis, *italics* for subtle highlighting,",
    "and `inline code` for technical terms like `add_viz()`.",
    "",
    "Combine styles: ***bold and italic*** or **`bold code`**.",
    "Add [hyperlinks](https://dashboardr.dev) to external resources.",
    "",
    "Lists work too:",
    "",
    "- First item with **bold**",
    "- Second item with *italics*",
    "- Third item with `code`"
  )) %>%
  add_accordion(
    title = "What is an accordion?",
    text = "An **accordion** is a collapsible content block. Users click to
           expand and reveal hidden content. Great for FAQs or code examples."
  ) %>%
  add_accordion(
    title = "Pro tip: Nested content",
    text = md_text(
      "Inside accordions, you can include:",
      "",
      "- Markdown formatting like **bold** and *italics*",
      "- Code blocks for examples",
      "- Links to documentation"
    )
  ) %>%
  add_text("Text-only pages are useful for documentation or methodology notes.")

# Create page and add content
text_page <- create_page(name = "Text & Content", icon = "ph:chalkboard-simple-bold") %>%
  add_content(text_content)
```

</details>


This page demonstrates text formatting and content blocks available in dashboardr.

You can use **bold text** for emphasis, *italics* for subtle highlighting, and `inline code` for technical terms or function names like `add_viz()`.

Combine styles: ***bold and italic*** or **`bold code`**. Add [hyperlinks](https://dashboardr.dev) to external resources.

Lists work too:

- First item with **bold**
- Second item with *italics*
- Third item with `code`


<details>
<summary>{{< iconify ph question >}} What is an accordion?</summary>

An **accordion** is a collapsible content block. Users click to expand and reveal hidden content. Great for FAQs, code examples, or additional details that might clutter the main page.

</details>


<details>
<summary>{{< iconify ph lightbulb >}} Pro tip: Nested content</summary>

Inside accordions, you can include:

- Markdown formatting like **bold** and *italics*
- Code blocks for examples
- Links to documentation

This keeps your dashboard clean while providing depth for curious users.

</details>


---

Text-only pages are useful for documentation, methodology notes, data sources, or any context that helps users understand your visualizations.

