---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE, warning=FALSE, message=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
library(tidyverse)
library(dashboardr)
```

# dashboardr <img src="man/figures/logo.svg" align="right" height="139" alt="" />

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/dashboardr)](https://CRAN.R-project.org/package=dashboardr)
<!-- badges: end -->

**dashboardr** makes it easy to create beautiful, interactive [Quarto dashboards](https://quarto.org/docs/dashboards/) in R with minimal code. Perfect for survey data, analytics reports, and data storytelling.

## Installation

```{r install, eval = F}
# install.packages("pak")
pak::pak("favstats/dashboardr")
```


## See some Example Dashboards

`dashboardr` includes two built-in demo dashboards that showcase its capabilities:

1. [**Tutorial Dashboard**](https://favstats.github.io/dashboardr/live-demos/tutorial/docs/index.html) - A beginner-friendly dashboard demonstrating basic features
2. [**Showcase Dashboard**](https://favstats.github.io/dashboardr/live-demos/showcase/docs/index.html) - A comprehensive dashboard showcasing advanced features

Both dashboards use real data from the General Social Survey (GSS) and can be generated with a single function call! 

## The Grammar of Dashboards

Just as ggplot2 gave us a **grammar of graphics**, dashboardr provides a **grammar of dashboards**. Instead of wrestling with configuration files and layout code, you compose dashboards from six intuitive building blocks:

**Data** + **Content** + **Visualizations** + **Layout** + **Navigation** + **Styling** = Dashboard

```{r quickstart}
library(dashboardr)

# 1. Data - What you're working with
data <- mtcars %>%
  mutate(
    cyl_label = paste(cyl, "cylinders"),
    am_label = ifelse(am == 0, "Automatic", "Manual")
  )

# 2. Content - Visualizations + text + images (unified system)
content <- create_viz(type = "histogram") %>%
  add_viz(x_var = "mpg", title = "MPG Distribution", tabgroup = "overview") %>%
  add_viz(x_var = "hp", title = "Horsepower", tabgroup = "overview") %>%
  add_text("## Key Insights", "", "The data shows interesting patterns...") %>%
  add_callout("Note: All visualizations are interactive!", type = "tip")

# 3. Dashboard - Layout + Navigation + Styling
dashboard <- create_dashboard(
  title = "Car Analysis Dashboard",
  output_dir = "my_dashboard",
  tabset_theme = "modern"  # Styling
) %>%
  add_page(  # Navigation
    "Analysis",
    data = data,
    content = content,  # Can use 'content' or 'visualizations' - they're interchangeable!
    is_landing_page = TRUE
  )

# Generate
generate_dashboard(dashboard)
```

This creates a complete Quarto dashboard with interactive visualizations!


## ✨ Features

`dashboardr` works with **pre-rendered static HTMLs** which in many cases makes it faster to load than a Shiny application (provided a good internet connection). It also is more robust since no new graphs have to be calculated.

- 🎨 **6 Built-in Tabset Themes** - Modern, minimal, pills, classic, underline, segmented
- 📊 **Rich Visualizations** - Histograms, timelines, stacked bars, heatmaps, bar charts
- 🎯 **Smart Defaults** - Set parameters once, reuse everywhere
- 🔍 **Advanced Filtering** - Row-level filters for each visualization
- 📚 **Multi-Dataset Support** - Work with multiple datasets seamlessly
- 🎭 **Nested Tabgroups** - Create hierarchical tab structures
- ⚡ **Vectorized Creation** - Generate multiple visualizations efficiently
- 🎬 **Loading Overlays** - 4 animated overlay themes
- 🧭 **Flexible Navigation** - Navbar menus, dropdown menus, icons
- 📱 **Responsive Design** - Works on all screen sizes



## Core Workflow: Data → Visualizations → Dashboard

Creating a dashboard follows a simple three-step pattern:

### Step 1: Build Visualizations

Use `create_viz()` to set shared defaults, then `add_viz()` for each chart:

```{r step1, eval=T}
library(dashboardr)

# Set defaults once (type, colors, etc.)
my_viz <- create_viz(
  type = "histogram",           # All charts will be histograms
  color_palette = c("#3498DB"),  # All charts use this color
  bins = 30                      # All charts use 30 bins
) %>%
  # Add individual visualizations
  add_viz(
    x_var = "age",              # What to plot
    title = "Age Distribution",  # Chart title
    tabgroup = "overview"        # Which tab group
  ) %>%
  add_viz(
    x_var = "income",
    title = "Income Distribution",
    tabgroup = "overview",
    bins = 50  # Override: this one uses 50 bins
  )

# See what you built
print(my_viz)
```

**Key concepts:**

- `create_viz()`: Sets defaults that apply to all visualizations
- `add_viz()`: Adds one visualization, can override any default
- `tabgroup`: Organizes visualizations into tabs (e.g., "overview", "demographics/age")
- `print()`: Shows the structure before generating

### Step 2: Build Dashboard Structure

Use `create_dashboard()` to configure, then `add_page()` for each page:


```{r step2, eval=T}
survey_data <- cars

dashboard <- create_dashboard(
  title = "Employee Survey Dashboard",
  output_dir = "my_first_dashboard", # where to put the qmd files
  tabset_theme = "modern"  # Tab styling
) %>%
  # Landing page with text
  add_page(
    "Home",
    text = md_text(
      "# Welcome!",
      "",
      "This dashboard presents employee survey results.",
      "",
      "Navigate using the tabs above to explore different analyses."
    ),
    is_landing_page = TRUE
  ) %>%
  # Analysis page with data + visualizations
  add_page(
    "Analysis",
    data = survey_data,        # Your data
    visualizations = my_viz,    # The viz you created above
    icon = "ph:chart-line"      # Optional icon
  )

# See the dashboard structure
print(dashboard)
```

**Key concepts:**

- `create_dashboard()`: Sets dashboard-level options (title, theme, output location)
- `add_page()`: Adds a page to the navbar
- `md_text()`: Creates markdown text blocks (headings, paragraphs, etc.)
- `data`: Attaches your dataset to a page (available to all visualizations on that page)
- `is_landing_page`: Makes this the default page users see first

### Step 3: Generate HTML

Use `generate_dashboard()` to create the actual dashboard:

```{r step3, eval=FALSE}
# Generate QMD files only (fast, for development)
generate_dashboard(dashboard, render = FALSE)

# Generate QMD files AND render to HTML (slower, for final output)
generate_dashboard(dashboard, render = TRUE)

# Generate + render + open in browser
generate_dashboard(dashboard, render = TRUE, open = "browser")
```

**Key concepts:**

- `render = FALSE`: Only creates Quarto files (.qmd), doesn't run Quarto
- `render = TRUE`: Creates files AND renders to HTML (requires Quarto CLI)
- `open = "browser"`: Opens the dashboard in your browser after rendering

### Complete Example

```{r complete, eval=FALSE}
library(dashboardr)
library(dplyr)

# Your data
data <- mtcars %>%
  mutate(cyl_label = paste(cyl, "cylinders"))

# Step 1: Visualizations
viz <- create_viz(type = "histogram") %>%
  add_viz(x_var = "mpg", title = "MPG", tabgroup = "overview") %>%
  add_viz(x_var = "hp", title = "Horsepower", tabgroup = "overview")

# Step 2: Dashboard
dashboard <- create_dashboard(
  title = "Car Dashboard",
  output_dir = "my_dashboard"
) %>%
  add_page("Home", text = md_text("# Welcome!"), is_landing_page = TRUE) %>%
  add_page("Charts", data = data, visualizations = viz)

# Step 3: Generate
generate_dashboard(dashboard, render = TRUE, open = "browser")
```

That's it! You now have a complete interactive dashboard.

### Composing Visualizations with `+`

You can combine visualization collections using the `+` operator (like ggplot2):

```{r compose, eval=FALSE}
# Create separate collections for different topics
demographics <- create_viz(type = "histogram") %>%
  add_viz(x_var = "age", title = "Age", tabgroup = "demographics")

feedback <- create_viz(type = "histogram") %>%
  add_viz(x_var = "satisfaction", title = "Satisfaction", tabgroup = "feedback")

# Combine them!
combined <- demographics + feedback

print(combined)
#> 📊 VISUALIZATION COLLECTION
#> Total visualizations: 2
#> STRUCTURE:
#> ├─ 📁 demographics
#> │  └─ 📉 HISTOGRAM: Age
#> └─ 📁 feedback
#>    └─ 📉 HISTOGRAM: Satisfaction
```

**When to use `+`:**

- Organize complex dashboards into logical modules
- Combine visualizations from different scripts/team members
- Keep related visualizations grouped together in your code

## Key Features

### 🎯 Smart Defaults

Set common parameters once and reuse them:

```{r defaults}
viz <- create_viz(
  type = "stackedbars",
  questions = paste0("Q", 1:5),
  stacked_type = "percent",
  horizontal = TRUE,
  color_palette = c("#E74C3C", "#95A5A6", "#27AE60")
) %>%
  add_viz(title = "Wave 1", filter = ~ wave == 1) %>%
  add_viz(title = "Wave 2", filter = ~ wave == 2) %>%
  add_viz(title = "Wave 3", filter = ~ wave == 3)
```

### ⚡ Vectorized Creation

Create multiple visualizations efficiently:

```{r vectorized}
viz <- create_viz(
  type = "bar",
  horizontal = TRUE,
  bar_type = "percent"
) %>%
  add_vizzes(
    x_var = paste0("question_", 1:10),
    title = paste("Question", 1:10),
    .tabgroup_template = "survey/{title}"
  )
# Creates 10 visualizations with one call!
```

### 🎭 Nested Tabgroups

Create hierarchical structures:

```{r nested}
viz <- create_viz(
  type = "stackedbar",
  x_var = "response",
  stack_var = "category"
) %>%
  # Wave 1 tabs
  add_viz(title = "Overall", tabgroup = "survey/wave1", filter = ~ wave == 1) %>%
  add_viz(title = "By Age", tabgroup = "survey/wave1/age", filter = ~ wave == 1) %>%
  add_viz(title = "By Gender", tabgroup = "survey/wave1/gender", filter = ~ wave == 1) %>%
  # Wave 2 tabs
  add_viz(title = "Overall", tabgroup = "survey/wave2", filter = ~ wave == 2) %>%
  add_viz(title = "By Age", tabgroup = "survey/wave2/age", filter = ~ wave == 2) %>%
  add_viz(title = "By Gender", tabgroup = "survey/wave2/gender", filter = ~ wave == 2) %>%
  # Custom labels
  set_tabgroup_labels(list(
    survey = "📊 Survey Results",
    wave1 = "Wave 1 (2024)",
    wave2 = "Wave 2 (2025)"
  ))
```

### 📚 Multi-Dataset Support

```{r multi-dataset, eval=F}
dashboard %>%
  add_page(
    "Analysis",
    data = list(
      sales = sales_data,
      customers = customer_data,
      products = product_data
    ),
    visualizations = viz
  )
```

### 🎬 Loading Overlays

```{r overlay, eval=FALSE}
dashboard %>%
  add_page(
    "Reports",
    data = large_dataset,
    visualizations = viz,
    overlay = TRUE,
    overlay_theme = "glass",
    overlay_text = "Loading reports..."
  )
```

## Visualization Types

| Function | Description | Use Case |
|----------|-------------|----------|
| `create_histogram()` | Distribution visualization | Age, income, scores |
| `create_bar()` | Grouped/clustered bars | Category comparisons |
| `create_stackedbar()` | Single stacked bar | Likert scales, compositions |
| `create_stackedbars()` | Multiple stacked bars | Multiple questions |
| `create_timeline()` | Time series | Trends over time |
| `create_heatmap()` | 2D heatmap | Correlations, matrices |

## Themes

Choose from 6 built-in themes:

- **modern** - Clean, centered tabs with subtle shadows
- **minimal** - Simple, flat design
- **pills** - Rounded pill-shaped tabs
- **classic** - Traditional tabbed interface
- **underline** - Underlined active tabs
- **segmented** - Segmented control style

```{r themes}
dashboard <- create_dashboard(
  title = "My Dashboard",
  output_dir = "output",
  tabset_theme = "modern",
  tabset_colors = list(
    active_bg = "#3498DB",
    active_text = "#FFFFFF"
  )
)
```

## Documentation

- **Getting Started**: `vignette("getting-started")`
- **Advanced Features**: `vignette("advanced-features")`
- **Specific Visualizations**:
  - `vignette("bar_vignette")`
  - `vignette("timeline_vignette")`
  - `vignette("stackedbar_vignette")`
  - `vignette("stackedbars_vignette")`
  - `vignette("heatmap_vignette")`


## Real-World Use Cases

- 🎓 **Academic Research** - Present study results interactively
- 📊 **Survey Analysis** - Visualize Likert-scale responses across waves
- 📈 **Business Analytics** - Track KPIs over time by department
- 💼 **Consulting Reports** - Create client-ready dashboards
- 📱 **Data Journalism** - Build interactive data stories

## Contributing

Contributions are welcome! Please feel free to submit a Pull Request.

## License

MIT License - see [LICENSE.md](LICENSE.md) for details.

## Citation

```{r citation, eval=T}
citation("dashboardr")
```
