---
title: "Getting Started With create_stackedbars()"
author: "Alexandra Pafford"
date: "2025-07-08"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette:
  %\VignetteIndexEntry{Stacked Bar Chart Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8"}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, # Default figure width
  fig.height = 5, # Default figure height
  warning = FALSE, # Suppress warnings from highcharter, haven etc. for cleaner output
  message = FALSE  # Suppress messages (e.g., haven conversion notes) for cleaner output
)
```

## Introduction

This vignette demonstrates how to use the
`create_stackedbar()` and `create_stackedbar_tabs()` functions to
produce interactive stacked-bar charts in HTML, using the `iris` data
set as a stand-in for survey results.

## 1. Prepare the Data

We’ll convert `iris` into a “survey” table by:

- Introducing a few `NA` values  
- Treating `Species` as a categorical question  
- Binning `Petal.Length` into categories  

```{r data-prep}
iris_data <- as_tibble(iris) %>%
  # Inject some NAs
  mutate(
    Species     = if_else(row_number() %% 10 == 0, NA_character_, as.character(Species)),
    Petal_Length = if_else(row_number() %% 15 == 0, NA_real_, Petal.Length)
  )

# Define binning for Petal.Length
petal_breaks <- c(0, 1.5, 3.0, 4.5, 6.0, Inf)
petal_labels <- c(
  "Very Short (0-1.5)",
  "Short (1.5-3.0)",
  "Medium (3.0-4.5)",
  "Long (4.5-6.0)",
  "Very Long (>6.0)"
)

# Desired display order
petal_order <- petal_labels

# A simple color palette
pal <- c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba", "gray")
```

## 2. Single Stacked-Bar Chart

Let’s see how `Petal.Length` categories distribute across `Species`.

```{r single-chart, echo=TRUE}
chart1 <- create_stackedbar(
  data               = iris_data,
  x_var              = "Species",
  stack_var          = "Petal_Length",
  title              = "Petal Length Distribution by Iris Species",
  subtitle           = "Including explicit (NA) category",
  x_label            = "Species",
  y_label            = "Count",
  stack_label        = "Petal Length Category",
  stacked_type       = "normal",
  include_na         = TRUE,
  stack_breaks       = petal_breaks,
  stack_bin_labels   = petal_labels,
  stack_order        = petal_order,
  color_palette      = pal
)

chart1
```

## 3. Multiple Charts in Tabs

Suppose we have several “survey items” (`Petal.Length`, `Petal.Width`,
`Sepal.Length`) and want to toggle between their distributions by
`Species`.

```{r multi-tabs, echo=TRUE}
stack_items <- c("Petal_Length", "Petal.Width", "Sepal.Length")
tab_labels  <- c("Petal Length",  "Petal Width",  "Sepal Length")

tabs <- create_stackedbar_tabs(
  data               = iris_data,
  x_var              = "Species",
  stack_vars         = stack_items,
  stack_var_labels   = tab_labels,
  common_title       = "Distribution of",
  subtitle           = "by Iris Species",
  x_label            = "Species",
  y_label            = "Percentage",
  stack_label        = "Category",
  stacked_type       = "percent",
  tooltip_suffix     = "%",
  include_na         = FALSE,
  # For demonstration we only bin Petal.Length; others use raw values
  stack_breaks       = petal_breaks,
  stack_bin_labels   = petal_labels,
  stack_order        = petal_order,
  color_palette      = pal
)

tabs
```

## Conclusion

The `create_stackedbars` function provides a powerful and flexible tool for visualizing data relationships, particularly useful for survey data. By leveraging its various arguments for mapping, binning, and ordering, you can transform raw, coded, or continuous data into clear, interactive, and professional-looking stacked bar charts with minimal effort. This makes it ideal for generating dashboards, allowing users to quickly grasp key insights from complex datasets.

You are easily able to flip through a battery of survey items to quickly compare responses to a single topic or idea, with the visualization flexibility of a typical stacked bar chart.

Experiment with different combinations of arguments and your own datasets to unlock the full potential of this function!