---
title: "{{< iconify ph users-three >}} Demographics"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

Explore how happiness, trust, and wellbeing vary across demographic groups.

Use the tabs below to examine different aspects of the data.

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

# Load data
data <- readRDS('dataset_2867obs.rds')

# Data summary
cat('Dataset loaded:', nrow(data), 'rows,', ncol(data), 'columns\n')

```

## Demographics & Education


::: {.panel-tabset}

### {{< iconify ph chart-bar >}} Happiness Distribution Across Education Levels


How happy are you with your life right now?

```{r demographics}
# Happiness Distribution Across Education Levels
result <- viz_stackedbar(
  data = data,
  title = "Happiness Distribution Across Education Levels",
  x_var = "degree_1a",
  stack_var = "happy_1a",
  subtitle = "Percentage breakdown within each education category",
  x_label = "Education Level",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 500)
}
result <- htmltools::div(
  style = 'height: 500px !important; min-height: 500px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "degree_1a",
        stack_var = "happy_1a",
        title = "Happiness Distribution Across Education Levels",
        subtitle = "Percentage breakdown within each education category",
        x_label = "Education Level",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "How happy are you with your life right now?",
        text_position = "above",
        icon = "ph:chart-bar",
        height = 500,
        tabgroup = "demographics")
```
[Full documentation](../articles/showcase_dashboard_code.html#demographics-happiness-education)

</details>

## Demographics & Education


::: {.panel-tabset}

### {{< iconify ph gender-intersex >}} Happiness Distribution by Gender

```{r demographics-2}
# Happiness Distribution by Gender
result <- viz_stackedbar(
  data = data,
  title = "Happiness Distribution by Gender",
  x_var = "sex_1a",
  stack_var = "happy_1a",
  subtitle = "Gender differences in reported happiness levels",
  x_label = "Gender",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 450)
}
result <- htmltools::div(
  style = 'height: 450px !important; min-height: 450px !important; width: 100%; overflow: visible;',
  result
)

result
```

How happy are you with your life right now?



:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "sex_1a",
        stack_var = "happy_1a",
        title = "Happiness Distribution by Gender",
        subtitle = "Gender differences in reported happiness levels",
        x_label = "Gender",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "How happy are you with your life right now?",
        text_position = "below",
        icon = "ph:gender-intersex",
        height = 450,
        tabgroup = "demographics")
```
[Full documentation](../articles/showcase_dashboard_code.html#demographics-happiness-gender)

</details>

## Political Attitudes


::: {.panel-tabset}

### {{< iconify ph users-three >}} Party ID by Political Views


This shows how party identification aligns with political ideology.

```{r politics}
# Party ID by Political Views
result <- viz_stackedbar(
  data = data,
  title = "Party ID by Political Views",
  x_var = "polviews_1a",
  stack_var = "partyid_1a",
  subtitle = "Distribution of party identification across political ideology",
  x_label = "Political Views",
  y_label = "Percentage of Respondents",
  stack_label = "Party ID",
  stacked_type = "percent",
  tooltip_suffix = "%",
  color_palette = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 550)
}
result <- htmltools::div(
  style = 'height: 550px !important; min-height: 550px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "polviews_1a",
        stack_var = "partyid_1a",
        title = "Party ID by Political Views",
        subtitle = "Distribution of party identification across political ideology",
        x_label = "Political Views",
        y_label = "Percentage of Respondents",
        stack_label = "Party ID",
        stacked_type = "percent",
        tooltip_suffix = "%",
        color_palette = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2"),
        text = "This shows how party identification aligns with political ideology.",
        text_position = "above",
        icon = "ph:users-three",
        height = 550,
        tabgroup = "politics")
```
[Full documentation](../articles/showcase_dashboard_code.html#politics-party-views)

</details>

## Political Attitudes


::: {.panel-tabset}

### {{< iconify ph map-pin >}} Trust Levels by US Region

```{r politics-2}
# Trust Levels by US Region
result <- viz_stackedbar(
  data = data,
  title = "Trust Levels by US Region",
  x_var = "region_1a",
  stack_var = "trust_1a",
  subtitle = "Regional variation in interpersonal trust",
  x_label = "US Region",
  y_label = "Percentage of Respondents",
  stack_label = "Trust Level",
  stack_order = c("Can Trust", "Can't Be Too Careful", "Depends"),
  stacked_type = "percent",
  tooltip_suffix = "%",
  color_palette = c("#2E8B57", "#DAA520", "#CD5C5C")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 500)
}
result <- htmltools::div(
  style = 'height: 500px !important; min-height: 500px !important; width: 100%; overflow: visible;',
  result
)

result
```

Do you think you can usually trust strangers?



:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "region_1a",
        stack_var = "trust_1a",
        title = "Trust Levels by US Region",
        subtitle = "Regional variation in interpersonal trust",
        x_label = "US Region",
        y_label = "Percentage of Respondents",
        stack_label = "Trust Level",
        stack_order = c("Can Trust", "Cant Be Too Careful", "Depends"),
        stacked_type = "percent",
        tooltip_suffix = "%",
        color_palette = c("#2E8B57", "#DAA520", "#CD5C5C"),
        text = "Do you think you can usually trust strangers?",
        text_position = "below",
        icon = "ph:map-pin",
        height = 500,
        tabgroup = "politics")
```
[Full documentation](../articles/showcase_dashboard_code.html#politics-trust-region)

</details>

## Political Attitudes


::: {.panel-tabset}

### {{< iconify ph chart-pie >}} Gender Distribution Across Social Classes


Gender distribution across different social class categories.

```{r politics-3}
# Gender Distribution Across Social Classes
result <- viz_stackedbar(
  data = data,
  title = "Gender Distribution Across Social Classes",
  x_var = "class_1a",
  stack_var = "sex_1a",
  subtitle = "With custom labels and ordering",
  x_label = "Self-Reported Social Class",
  y_label = "Number of Respondents",
  stack_label = "Gender",
  x_order = c("Lower Class", "Working Class", "Middle Class", "Upper Class"),
  stack_order = c("Female", "Male"),
  stacked_type = "counts",
  tooltip_prefix = "Count: ",
  color_palette = c("#E07A5F", "#3D5A80")
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 450)
}
result <- htmltools::div(
  style = 'height: 450px !important; min-height: 450px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "stackedbar",
        x_var = "class_1a",
        stack_var = "sex_1a",
        title = "Gender Distribution Across Social Classes",
        subtitle = "With custom labels and ordering",
        x_label = "Self-Reported Social Class",
        y_label = "Number of Respondents",
        stack_label = "Gender",
        x_order = c("Lower Class", "Working Class", "Middle Class", "Upper Class"),
        stack_order = c("Female", "Male"),
        stacked_type = "counts",
        tooltip_prefix = "Count: ",
        color_palette = c("#E07A5F", "#3D5A80"),
        text = "Gender distribution across different social class categories.",
        text_position = "above",
        icon = "ph:chart-pie",
        height = 450,
        tabgroup = "politics")
```
[Full documentation](../articles/showcase_dashboard_code.html#politics-gender-class)

</details>

## Social Issues


::: {.panel-tabset}

### {{< iconify ph heatmap >}} Trust by Education and Age

```{r social}
# Trust by Education and Age
result <- viz_heatmap(
  data = data,
  title = "Trust by Education and Age",
  x_var = "degree_1a",
  y_var = "age_1a",
  value_var = "trust_1a",
  subtitle = "Average trust levels across education and age groups",
  x_label = "Education Level",
  y_label = "Age Group",
  value_label = "Trust Level",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 600)
}
result <- htmltools::div(
  style = 'height: 600px !important; min-height: 600px !important; width: 100%; overflow: visible;',
  result
)

result
```

This heatmap reveals trust patterns across education and age groups.



:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "heatmap",
        x_var = "degree_1a",
        y_var = "age_1a",
        value_var = "trust_1a",
        title = "Trust by Education and Age",
        subtitle = "Average trust levels across education and age groups",
        x_label = "Education Level",
        y_label = "Age Group",
        value_label = "Trust Level",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "This heatmap reveals trust patterns across education and age groups.",
        text_position = "below",
        icon = "ph:heatmap",
        height = 600,
        tabgroup = "social")
```
[Full documentation](../articles/showcase_dashboard_code.html#social-trust-education-age)

</details>

## Social Issues


::: {.panel-tabset}

### {{< iconify ph chart-pie >}} Trust by Region and Education


Educational and regional patterns in trust distribution.

```{r social-2}
# Trust by Region and Education
result <- viz_heatmap(
  data = data,
  title = "Trust by Region and Education",
  x_var = "region_1a",
  y_var = "degree_1a",
  value_var = "trust_1a",
  subtitle = "Educational and regional patterns in trust levels",
  x_label = "Region",
  y_label = "Education Level",
  value_label = "Trust Level",
  y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Force container height with explicit wrapper
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_size(result, height = 550)
}
result <- htmltools::div(
  style = 'height: 550px !important; min-height: 550px !important; width: 100%; overflow: visible;',
  result
)

result
```


:::


<details>
<summary>{{< iconify ph code >}} View R Code</summary>

```r
add_viz(type = "heatmap",
        x_var = "region_1a",
        y_var = "degree_1a",
        value_var = "trust_1a",
        title = "Trust by Region and Education",
        subtitle = "Educational and regional patterns in trust levels",
        x_label = "Region",
        y_label = "Education Level",
        value_label = "Trust Level",
        y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "Educational and regional patterns in trust distribution.",
        text_position = "above",
        icon = "ph:chart-pie",
        height = 550,
        tabgroup = "social")
```
[Full documentation](../articles/showcase_dashboard_code.html#social-trust-region-education)

</details>

