---
title: "{{< iconify ph info >}} About"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

# New Features Showcase

This dashboard demonstrates every new feature added in dashboardr v0.3.

## New Visualization Types

| Type | Function | Description |
|------|----------|-------------|
| Pie | `viz_pie()` | Classic pie chart with labels and tooltips |
| Donut | `viz_pie()` via `type = "donut"` | Pie with inner cut-out (auto 50%) |
| Lollipop | `viz_lollipop()` | Stem + dot, a clean bar chart alternative |
| Dumbbell | `viz_dumbbell()` | Two-point comparison per category |
| Gauge | `viz_gauge()` | KPI speedometer with color bands |
| Funnel | `viz_funnel()` | Sequential stage drop-off |
| Pyramid | `viz_funnel()` via `type = "pyramid"` | Reversed funnel |
| Sankey | `viz_sankey()` | Flow diagram between nodes |
| Waffle | `viz_waffle()` | Grid-of-squares proportional chart |

## Chart Export

Every chart on this dashboard has an export button (top-right hamburger menu).
Supported formats: PNG, SVG, PDF, CSV, XLS, and fullscreen mode.

Set per-chart with `export = TRUE` or dashboard-wide with `chart_export = TRUE`.

## Reference Lines & Annotations

Timeline charts now support:

- `reference_lines`: horizontal/vertical guidelines (e.g. targets, thresholds)
- `annotations`: labeled markers for notable events

## show_when Comparison Operators

The conditional visibility system now supports:

- `~ score > 50` (greater than)
- `~ score < 100` (less than)
- `~ year >= 2020` (greater than or equal)
- `~ score <= 75` (less than or equal)
- `~ !active` (negation)

Combined with the existing `==`, `!=`, `%in%`, `&`, and `|` operators,
this makes `show_when` a complete conditional logic system.

## Built With

- [dashboardr](https://github.com/favstats/dashboardr) - Static HTML dashboards in R
- [Highcharts](https://www.highcharts.com/) via highcharter - Interactive charts
- [Quarto](https://quarto.org/) - Document rendering engine

```{r, echo=FALSE, results='asis'}
# Enable chart export buttons (download as PNG/SVG/PDF/CSV)
dashboardr::enable_chart_export()
```

