---
title: "Test Layout"
format: html
---

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
dashboardr::.page_config(accessibility = TRUE)
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


# Trying to obtain this:


### A row with multiple objects / metrics / custom html elements


```{r}
#| echo: false
dashboardr::render_value_box_row(list(
  list(
    title = "VB 1",
    value = 100,
    bg_color = "#cb333b",
    logo_url = NULL,
    logo_text = NULL
  ),
  list(
    title = "VB 2",
    value = 200,
    bg_color = "#c6d7ad",
    logo_url = NULL,
    logo_text = NULL
  ),
  list(
    title = "VB 3",
    value = 300,
    bg_color = "#2c3e50",
    logo_url = NULL,
    logo_text = NULL
  )
))
```


::: {.callout-note}
Add_value_box_row works correctly with multiple value boxes
:::


### Trying to obtain the same result using

- **add_layout_column**
- **add_layout_row**


::: {.callout-warning}
The layout functions are not working as expected, the metrics are not being displayed in the same row and column as they should be,
              and the metric element do not properly render
:::


::: {layout-ncol=2}

```{r}
#| echo: false
dashboardr::html_metric(value = 100, title = "metric 1 - this is between add_layout_row and column", icon = NULL, color = "#cb333b", subtitle = "subtitle1", aria_label = NULL)
```


```{r}
#| echo: false
dashboardr::html_metric(value = 2000, title = "metric 2 -this is also between add_layout_row and column", icon = NULL, color = NULL, subtitle = "subtitle2", aria_label = NULL)
```

:::


```{r}
#| echo: false
dashboardr::html_metric(value = 3000, title = "metric 3 - this is alone", icon = NULL, color = NULL, subtitle = "subtitle2", aria_label = NULL)
```

