  ---
title: "Live Demos: Tutorial & Showcase Dashboards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Live Demos: Tutorial & Showcase Dashboards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`dashboardr` includes several live demo dashboards that showcase its capabilities:

| Demo | Description | Link |
|------|-------------|------|
| **Features Demo** | Comprehensive showcase of tab styling, inputs, and overlays | [View Demo](../live-demos/features/index.html) |
| **Tutorial Dashboard** | Beginner-friendly dashboard demonstrating basic features | [View Demo](../live-demos/tutorial/index.html) |
| **Showcase Dashboard** | Advanced dashboard with multiple pages and tabsets | [View Demo](../live-demos/showcase/index.html) |

All demos use real data from the General Social Survey (GSS) and can be generated with a single function call!

## Features Demo (Recommended)

The **Features Demo** is the best place to explore dashboardr's interactive capabilities:

- **Tab Styling** - See different `tabset_theme` options (modern, pills) and custom `tabset_colors`
- **Interactive Inputs** - Try select dropdowns, checkboxes, sliders, and radio buttons
- **Loading Overlays** - See the glass overlay effect (reload the page!)
- **Page Alignment** - Notice the "About" page aligned to the right

### What Generates It

```r
# Tab styling page
create_page("Corporate Colors", data = gss, type = "bar",
  tabset_theme = "pills",
  tabset_colors = list(
    active_bg = "#2C3E50",
    active_text = "#FFFFFF",
    inactive_bg = "#ECF0F1"
  )
) %>%
  add_viz(x_var = "degree", title = "Education", tabgroup = "Section A")

# Interactive inputs page
create_page("Interactive Inputs", data = gss, type = "bar") %>%
  add_input_row() %>%
    add_input(input_id = "edu", label = "Education", 
              type = "select_multiple", filter_var = "degree") %>%
  end_input_row() %>%
  add_viz(x_var = "happy", title = "Happiness")

# Loading overlay page
create_page("Loading Overlay", data = gss, type = "bar",
  overlay = TRUE,
  overlay_theme = "glass",
  overlay_text = "Loading visualizations...",
  overlay_duration = 2000
) %>%
  add_viz(x_var = "degree", title = "Education")
```

## Tutorial Dashboard

The tutorial dashboard is perfect for:

- Learning the basics of dashboardr
- Understanding how tabsets work
- Seeing simple visualizations in action
- Getting started quickly

### Features Demonstrated

- **Stacked bar charts** with custom colors and ordering
- **Heatmaps** showing relationships between variables
- **Tabset grouping** for organizing related visualizations
- **Standalone charts** without tabsets
- **Text-only pages** for context and documentation
- **Icons** throughout the interface

### Running the Tutorial Dashboard

```{r setup, message=FALSE}
library(dashboardr)
```

```{r}
# Requires the 'gssr' package
# install.packages("gssr")

# Generate and open the tutorial dashboard
tutorial_dashboard()
```

This will:

1. Load GSS panel data
2. Create visualizations with stacked bars and heatmaps
3. Build a multi-page dashboard
4. Render it with Quarto
5. Open it in your browser

**Output directory:** `tutorial_dashboard/`

### What You'll See

The tutorial dashboard includes:

- **Welcome Page** - Introduction and navigation guide
- **Example Dashboard** - Two tabsets demonstrating different chart types
- **Standalone Charts** - Examples without tabsets
- **Text-Only Page** - Pure markdown content
- **Showcase Link** - Link to the more advanced showcase

### Code Example

Here's what the tutorial dashboard does internally:

```{r}
# Create visualization collection with tabsets
analysis_vizzes <- create_viz() %>%
  add_viz(
    type = "stackedbar",
    x_var = "degree_1a",
    stack_var = "happy_1a",
    title = "Happiness by Education",
    tabgroup = "demographics",  # First tabset
    stacked_type = "percent"
  ) %>%
  add_viz(
    type = "heatmap",
    x_var = "degree_1a",
    y_var = "age_1a",
    value_var = "trust_1a",
    title = "Trust Patterns",
    tabgroup = "social",  # Second tabset
  )

# Build dashboard
dashboard <- create_dashboard(
  title = "Tutorial Dashboard",
  output_dir = "tutorial_dashboard"
) %>%
  add_page(
    "Welcome",
    text = "Introduction text here",
    is_landing_page = TRUE
  ) %>%
  add_page(
    "Analysis",
    data = gss_data,
    visualizations = analysis_vizzes
  )

# Generate
generate_dashboard(dashboard, render = TRUE)
```

---

## Showcase Dashboard

The showcase dashboard is perfect for:

- Seeing the full power of dashboardr
- Understanding advanced features
- Getting inspiration for your own dashboards
- Demonstrating to stakeholders

### Features Demonstrated

Everything from the tutorial dashboard, plus:

- **Multiple tabset groups** (Demographics, Politics, Social Issues)
- **Complex visualizations** with custom styling
- **Mixed content pages** (text + visualizations)
- **Card layouts** with images and custom content
- **Advanced Quarto features** (breadcrumbs, search, code folding)
- **All visualization types** in one dashboard

### Running the Showcase Dashboard

```{r}
# Requires the 'gssr' package
# install.packages("gssr")

# Generate and open the showcase dashboard
showcase_dashboard()
```

This will:

1. Load GSS panel data
2. Create comprehensive visualizations across multiple tabsets
3. Build a multi-page dashboard with advanced features
4. Render it with Quarto
5. Open it in your browser

**Output directory:** `comprehensive_dashboard_test/`

### What You'll See

The showcase dashboard includes:

- **Welcome Page** - Feature overview
- **GSS Data Analysis** - Three tabsets:
  - Demographics & Education (2 visualizations)
  - Political Attitudes (3 visualizations)
  - Social Issues (2 visualizations)
- **Key Findings** - Mixed text and visualizations
- **Summary Charts** - Standalone charts highlighting key insights
- **About Page** - Card layout showcasing team members (with cat and dog photos!)

### Advanced Features

The showcase demonstrates:

```{r}
# Multiple tabset groups
analysis_viz <- create_viz() %>%
  add_viz(..., tabgroup = "demographics") %>%
  add_viz(..., tabgroup = "demographics") %>%
  add_viz(..., tabgroup = "politics") %>%
  add_viz(..., tabgroup = "politics") %>%
  add_viz(..., tabgroup = "social") %>%
  set_tabgroup_labels(list(
    demographics = "Demographics & Education",
    politics = "Political Attitudes",
    social = "Social Issues"
  ))

# Card layouts
card(
  content = "Team member bio",
  title = "Mario il Gatto",
  image = "https://images.unsplash.com/photo-cat",
  footer = "Website: mario-il-gatto.data"
)

# Mixed content pages
add_page(
  "Findings",
  text = md_text("Analysis summary..."),
  data = data,
  visualizations = viz
)
```

---

## Comparison

| Feature | Tutorial | Showcase |
|---------|----------|----------|
| **Pages** | 4 | 5 |
| **Tabsets** | 2 | 3 |
| **Visualizations** | 6 | 9 |
| **Chart Types** | Stackedbar, Heatmap | Stackedbar, Heatmap |
| **Standalone Charts** | ✅ Yes | ✅ Yes |
| **Card Layouts** | ❌ No | ✅ Yes |
| **Mixed Content** | ❌ No | ✅ Yes |
| **Complexity** | Beginner | Advanced |
| **Purpose** | Learning | Inspiration |

---

## Customizing the Demos

Both demo functions create dashboards in your working directory. You can:

1. **Run the demo** to see the output
2. **Inspect the generated files** in the output directory
3. **Modify the QMD files** if desired
4. **Use the code as a template** for your own dashboards

### Extracting the Code

Want to see the full code? Both functions are fully documented:

```{r}
# View the tutorial dashboard code
?tutorial_dashboard

# View the showcase dashboard code
?showcase_dashboard

# See the actual implementation
View(tutorial_dashboard)
View(showcase_dashboard)
```

---

## Requirements

Both demos require:

1. **Quarto CLI** installed on your system
2. **gssr package** for GSS data:
   ```{r}
   install.packages("gssr")
   ```

If you don't have these, you'll get helpful error messages guiding you to install them.

---

## Next Steps

After exploring the demos:

1. **Start with the tutorial** - Run `tutorial_dashboard()` to see basic features
2. **Try the showcase** - Run `showcase_dashboard()` to see advanced capabilities
3. **Read the vignettes** - Check out `vignette("getting-started")` for detailed guides
4. **Build your own** - Use the demos as templates for your projects!

---

## Tips for Using the Demos

### For Presentations

```{r}
# Generate without opening (for presentations)
dashboard <- tutorial_dashboard()

# Then open manually when ready
browseURL("tutorial_dashboard/docs/index.html")
```

### For Learning

```{r}
# Generate the tutorial
tutorial_dashboard()

# Explore the output directory
list.files("tutorial_dashboard", recursive = TRUE)

# Examine the generated QMD files
file.edit("tutorial_dashboard/index.qmd")
```

### For Inspiration

```{r}
# Generate the showcase
showcase_dashboard()

# Compare different visualizations
# Notice how tabgroups organize content
# See how standalone charts work
# Explore card layouts and mixed content
```

---

## Troubleshooting

### Quarto Not Found

```
Error: Quarto CLI not found
```

**Solution:** Install Quarto from https://quarto.org/docs/get-started/

### gssr Package Missing

```
Error: package 'gssr' not found
```

**Solution:**
```{r}
install.packages("gssr")
```

### Port Already in Use

If the dashboard doesn't open, the port might be busy. Try:

```{r}
# Generate without opening
tutorial_dashboard()

# Then open manually
browseURL("tutorial_dashboard/docs/index.html")
```

---

## Getting Help

```{r}
# Function documentation
?tutorial_dashboard
?showcase_dashboard

# Package overview
help(package = "dashboardr")

# All vignettes
vignette(package = "dashboardr")
```

---

**Happy dashboard building!** 🎉

Remember: The best way to learn dashboardr is to see it in action. Run both demos and explore the generated files!

