---
title: "Live Demos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Live Demos}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`dashboardr` includes several live demo dashboards that showcase its capabilities.

### Main Demos

- **Tutorial Dashboard** - Beginner-friendly introduction to dashboardr. [View Demo](../live-demos/tutorial/index.html)
- **Showcase Dashboard** - Advanced dashboard with multiple pages and tabsets. [View Demo](../live-demos/showcase/index.html)
- **Interactive Inputs** - Gallery of input types (select, slider, switch, checkbox). [View Demo](../live-demos/inputs/index.html)

### Tabset Theme Demos

Each tabset theme has its own demo dashboard showing 1, 2, and 3-level nested tabs:

- **Pills** - Rounded pill-shaped tabs. [View](../live-demos/tabsets/pills/index.html)
- **Modern** - Clean contemporary styling. [View](../live-demos/tabsets/modern/index.html)
- **Minimal** - Subtle, understated tabs. [View](../live-demos/tabsets/minimal/index.html)
- **Classic** - Traditional tab appearance. [View](../live-demos/tabsets/classic/index.html)
- **Underline** - Underlined active tab. [View](../live-demos/tabsets/underline/index.html)
- **Segmented** - iOS-style segmented controls. [View](../live-demos/tabsets/segmented/index.html)

### Loading Overlays

Demo of animated loading overlays with different themes (light, dark, glass, accent). [View Demo](../live-demos/overlay/index.html)

## Tutorial Dashboard

The tutorial dashboard is perfect for learning the basics:

- Stacked bar charts with custom colors and ordering
- Heatmaps showing relationships between variables
- Tabset grouping for organizing visualizations
- Text-only pages for documentation

```{r}
library(dashboardr)

# Generate and open the tutorial dashboard
tutorial_dashboard()
```

## Showcase Dashboard

The showcase dashboard demonstrates advanced features:

- Multiple tabset groups (Demographics, Politics, Social Issues)
- Complex visualizations with custom styling
- Mixed content pages (text + visualizations)
- Card layouts with images

```{r}
# Generate and open the showcase dashboard
showcase_dashboard()
```

## Setting a Tabset Theme

```{r}
# Set theme at dashboard level
create_dashboard(
  title = "My Dashboard",
  output_dir = "my_dashboard",
  tabset_theme = "pills"
)
```

Available themes: `pills`, `modern`, `minimal`, `classic`, `underline`, `segmented`.

## Creating Nested Tabs

```{r}
# 1 Level - Different tabgroup names create separate tabs
create_viz() %>%
  add_viz(type = "bar", x_var = "age", tabgroup = "age") %>%
  add_viz(type = "bar", x_var = "education", tabgroup = "education") %>%
  add_viz(type = "bar", x_var = "region", tabgroup = "region")

# 2 Levels - Use "/" to create parent > child hierarchy
create_viz() %>%
  add_viz(type = "bar", x_var = "age", tabgroup = "satisfaction/by_age") %>%
  add_viz(type = "bar", x_var = "education", tabgroup = "satisfaction/by_education")

# 3 Levels - Add more "/" for deeper nesting
create_viz() %>%
  add_viz(type = "bar", x_var = "age", tabgroup = "survey/satisfaction/age") %>%
  add_viz(type = "bar", x_var = "education", tabgroup = "survey/demographics/education")
```

## Loading Overlays

```{r}
# Add overlay to a page
add_page(
  name = "Analysis",
  data = my_data,
  visualizations = my_viz,
  overlay = TRUE,
  overlay_theme = "glass",
  overlay_text = "Loading charts...",
  overlay_duration = 2000
)
```

Available overlay themes: `light`, `dark`, `glass`, `accent`.

## Requirements

All demos require:

1. **Quarto CLI** installed on your system
2. **gssr package** for GSS data (for tutorial/showcase):

```{r}
install.packages("gssr")
```

## Next Steps

After exploring the demos:

1. Run `tutorial_dashboard()` to see basic features
2. Run `showcase_dashboard()` to see advanced capabilities
3. Check out `vignette("getting-started")` for detailed guides
4. Use the demos as templates for your own projects
