---
title: "Getting Started with dashboardr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with dashboardr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`dashboardr` is an R package for creating beautiful, interactive Quarto dashboards with minimal code. It provides a streamlined workflow for building dashboards with visualizations, custom layouts, and professional styling.

## Installation

```{r}
# Install from GitHub
# devtools::install_github("yourusername/dashboardr")

library(dashboardr)
library(dplyr)
```

## Quick Start

Creating a dashboard with `dashboardr` involves three main steps:

1. **Create** a dashboard project
2. **Add** pages and visualizations
3. **Generate** the Quarto dashboard

### Your First Dashboard

```{r}
# Prepare some data
my_data <- data.frame(
  category = rep(c("A", "B", "C"), each = 100),
  value = rnorm(300),
  score = sample(1:5, 300, replace = TRUE)
)

# Create visualizations
my_viz <- create_viz(
  type = "histogram",
  x_var = "value"
) %>%
  add_viz(title = "Distribution", tabgroup = "analysis")

# Create dashboard
dashboard <- create_dashboard(
  title = "My First Dashboard",
  output_dir = "my_dashboard"
) %>%
  add_page(
    name = "Home",
    data = my_data,
    visualizations = my_viz,
    is_landing_page = TRUE
  )

# Generate
generate_dashboard(dashboard)
```

This creates a complete Quarto dashboard with:
- Professional styling
- Interactive visualizations
- Responsive layout
- Modern tabset themes

## Core Concepts

### 1. Visualization Collections

Visualizations are created using `create_viz()` and `add_viz()`:

```{r}
viz <- create_viz() %>%
  add_viz(
    type = "histogram",
    x_var = "age",
    title = "Age Distribution"
  ) %>%
  add_viz(
    type = "stackedbar",
    x_var = "question",
    stack_var = "response",
    title = "Survey Responses"
  )
```

### 2. Dashboard Pages

Pages are added with `add_page()`:

```{r}
dashboard <- create_dashboard(
  title = "My Dashboard",
  output_dir = "output"
) %>%
  add_page("Home", text = "Welcome!", is_landing_page = TRUE) %>%
  add_page("Analysis", data = my_data, visualizations = viz) %>%
  add_page("About", text = "Dashboard created with dashboardr")
```

### 3. Dashboard Generation

Generate your dashboard with:

```{r}
generate_dashboard(dashboard)
```

This creates:
- QMD files for each page
- `_quarto.yml` configuration
- Data files (RDS format)
- Custom themes and styling

## Visualization Types

### Histogram

```{r}
create_viz(type = "histogram", x_var = "value") %>%
  add_viz(title = "Distribution", bins = 30)
```

### Stacked Bar

```{r}
create_viz(
  type = "stackedbar",
  x_var = "question",
  stack_var = "response"
) %>%
  add_viz(title = "Survey Results", horizontal = TRUE)
```

### Timeline

```{r}
create_viz(
  type = "timeline",
  time_var = "year",
  response_var = "score"
) %>%
  add_viz(title = "Trends Over Time")
```

### Bar Chart

```{r}
create_viz(
  type = "bar",
  x_var = "category",
  group_var = "segment"
) %>%
  add_viz(title = "Category Comparison")
```

## Styling and Themes

### Tabset Themes

Choose from 6 built-in themes:

```{r}
dashboard <- create_dashboard(
  title = "Styled Dashboard",
  output_dir = "styled",
  tabset_theme = "modern"  # or: minimal, pills, classic, underline, segmented
)
```

### Custom Colors

```{r}
dashboard <- create_dashboard(
  title = "Custom Colors",
  output_dir = "custom",
  tabset_theme = "modern",
  tabset_colors = list(
    active_bg = "#3498DB",
    active_text = "#FFFFFF"
  )
)
```

## Navigation Features

### Navbar Alignment

```{r}
dashboard %>%
  add_page("Home", text = "Home", is_landing_page = TRUE) %>%
  add_page("About", text = "About", navbar_align = "right")
```

### Dropdown Menus

```{r}
dashboard$navbar_sections <- list(
  navbar_menu(
    text = "Reports",
    pages = c("Monthly", "Quarterly", "Annual")
  )
)
```

### Icons

```{r}
add_viz(
  title = "Analysis",
  icon = "ph:chart-line",
  tabgroup = "reports"
)
```

## Loading Overlay

Add a loading animation to pages:

```{r}
add_page(
  "Analysis",
  data = large_dataset,
  visualizations = viz,
  overlay = TRUE,
  overlay_theme = "glass",
  overlay_text = "Loading analysis..."
)
```

## Performance Tips

### Quick Testing with Preview Mode

When developing dashboards with multiple pages, you don't need to regenerate everything to test changes:

```{r}
# Only generate the page you're working on
generate_dashboard(dashboard, preview = "Analysis")

# Test multiple specific pages
generate_dashboard(dashboard, preview = c("Home", "Analysis"))
```

**Why use it:**
- **10-50x faster** than full dashboard generation
- Perfect for iterative development
- Test visualization tweaks instantly
- Debug specific pages without waiting

**Example workflow:**
```{r}
# 1. Create dashboard with many pages
dashboard <- create_dashboard(...) %>%
  add_page("Home", ...) %>%
  add_page("Analysis", ...) %>%
  add_page("Demographics", ...) %>%
  add_page("Trends", ...) %>%
  add_page("About", ...)

# 2. Work on Analysis page
# Make changes to visualizations...

# 3. Quick test (< 1 second instead of 5+ seconds!)
generate_dashboard(dashboard, preview = "Analysis")

# 4. Once satisfied, generate full dashboard
generate_dashboard(dashboard, render = TRUE)
```

### Faster Subsequent Builds

Enable incremental builds to skip unchanged pages:

```{r}
# First build - generates all pages
generate_dashboard(dashboard, incremental = TRUE)

# Make changes to one page
# Second build - only regenerates that page!
generate_dashboard(dashboard, incremental = TRUE)
```

**How it works:**
- Automatically tracks which pages have changed
- Skips unchanged pages on subsequent builds
- Detects changes to data, visualizations, text, and parameters

**Perfect for:**
- Large dashboards (10+ pages)
- Iterative refinement
- Working with slow computations
- Development workflows

### Combine Both for Maximum Speed

```{r}
# Ultimate speed: preview + incremental
generate_dashboard(dashboard, 
                  preview = "Analysis",
                  incremental = TRUE)
```

**Result:** Sub-second build times! 🚀

## Troubleshooting

### Helpful Error Messages

`dashboardr` provides clear error messages with suggestions when things go wrong:

#### Typo in Visualization Type

```{r}
# Oops, typo!
add_viz(type = "histogra", x_var = "age")

# Error: Unknown type 'histogra'
# ℹ Did you mean 'histogram'?
# ℹ Available: histogram, bar, stackedbar, stackedbars, timeline, heatmap
```

#### Missing Required Parameter

```{r}
# Forgot x_var
create_histogram(data, title = "My Chart")

# Error: 'x_var' parameter is required
# ℹ Example: create_histogram(data, x_var = "age")
```

#### Typo in Page Name

```{r}
# Typo in preview
generate_dashboard(dashboard, preview = "Analisys")

# Error: Page 'analisys' not found
# ℹ Did you mean 'Analysis'?
# Available pages: Home, Analysis, About
```

#### Typo in Theme

```{r}
# Typo in theme name
create_dashboard(..., tabset_theme = "modrn")

# Error: Unknown theme 'modrn'
# ℹ Did you mean 'modern'?
# ℹ Available: modern, minimal, pills, classic, underline, segmented
```

### Common Issues and Solutions

#### Issue: Visualization Not Showing

**Problem:** Chart appears empty or doesn't render

**Solutions:**
```{r}
# 1. Check variable names match your data
names(my_data)  # Verify column names

# 2. Ensure data has rows
nrow(my_data)

# 3. Check for NA values
summary(my_data$x_var)

# 4. Use drop_na_vars to remove NAs
add_viz(
  type = "histogram",
  x_var = "age",
  drop_na_vars = TRUE  # Automatically removes NA rows
)
```

#### Issue: Page Not Found

**Problem:** Error when generating specific page

**Solutions:**
```{r}
# Check exact page names (case-sensitive in code, but preview is case-insensitive)
names(dashboard$pages)

# Use exact name in preview
generate_dashboard(dashboard, preview = "Analysis")  # Works
generate_dashboard(dashboard, preview = "analysis")  # Also works!
```

#### Issue: Slow Dashboard Generation

**Problem:** Takes a long time to generate

**Solutions:**
```{r}
# 1. Use preview mode during development
generate_dashboard(dashboard, preview = "WorkingPage")

# 2. Enable incremental builds
generate_dashboard(dashboard, incremental = TRUE)

# 3. Check data size
object.size(my_data)  # If > 10MB, consider filtering

# 4. Reduce visualization complexity
# - Fewer bins in histograms
# - Sample large datasets
# - Use filters to reduce data per viz
```

#### Issue: Tabs Not Organizing Correctly

**Problem:** Tabs appear in wrong order or nesting

**Solutions:**
```{r}
# 1. Check tabgroup syntax
add_viz(
  type = "histogram",
  x_var = "age",
  tabgroup = "demographics/age"  # Slash notation for nesting
)

# 2. Set explicit labels
viz %>% set_tabgroup_labels(list(
  demographics = "📊 Demographics",
  age = "Age Groups"
))

# 3. Use .insertion_index to control order (automatic)
# Tabs appear in the order you call add_viz()
```

### Getting More Help

```{r}
# View function documentation
?create_dashboard
?add_viz
?generate_dashboard

# See advanced examples
vignette("advanced-features")

# Check all available vignettes
vignette(package = "dashboardr")
```

## Next Steps

- **Advanced Features**: See `vignette("advanced-features")` for defaults, filters, nested tabs, and performance optimization
- **Visualization Details**: See individual vignettes for each viz type
- **Publishing**: See `?publish_dashboard` for deployment options

## Resources

- GitHub: [your-repo-url]
- Documentation: Run `?create_dashboard` for function help
- Examples: Check the `demo_*.R` files in the package

## Getting Help

```{r}
# Function documentation
?create_dashboard
?add_viz
?generate_dashboard

# Package overview
help(package = "dashboardr")
```

