---
title: "Creating Maps with viz_map()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Maps with viz_map()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## 📖 Introduction

The `viz_map()` function creates choropleth maps - geographic maps where regions are colored by data values. They're perfect for showing how metrics vary across countries, states, or other geographic areas.

```{r setup}
library(dashboardr)
library(dplyr)
```

```{r, include=FALSE}
devtools::load_all()
```

**Important**: Maps require data with geographic identifiers (ISO country codes, state abbreviations, etc.). Unlike other visualizations, maps typically can't use GSS data directly since GSS regions don't map to standard geographic codes. The examples below use hypothetical geographic datasets.

## 🌍 Basic World Map

### Country-Level Data

Your data needs a column with geographic codes that match the map:

```{r}
# Sample country data
country_data <- data.frame(
  country_code = c("USA", "GBR", "DEU", "FRA", "JPN", "CHN", "BRA", "AUS"),
  happiness_score = c(7.0, 7.2, 6.9, 6.5, 5.9, 5.1, 6.3, 7.3)
)

plot <- viz_map(
  data = country_data,
  join_var = "country_code",
  value_var = "happiness_score",
  map_type = "custom/world",
  title = "Happiness Score by Country"
)

plot
```

### Color Gradients

Control the color scale with `color_palette` - provide low and high colors:

```{r}
plot <- viz_map(
  data = country_data,
  join_var = "country_code",
  value_var = "happiness_score",
  map_type = "custom/world",
  title = "Happiness Score",
  value_label = "Score",
  color_palette = c("#FEE5D9", "#A50F15")  # Light to dark red
)

plot
```

## 🗺️ US State Maps

### State-Level Data

Use state abbreviations or FIPS codes:

```{r}
# State data with abbreviations
state_data <- data.frame(
  state = c("CA", "TX", "FL", "NY", "PA", "IL", "OH", "GA", "NC", "MI"),
  population_millions = c(39.5, 29.1, 21.5, 20.2, 13.0, 12.8, 11.8, 10.7, 10.4, 10.0)
)

plot <- viz_map(
  data = state_data,
  join_var = "state",
  value_var = "population_millions",
  map_type = "countries/us/us-all",
  title = "US Population by State",
  color_palette = c("#EBF5FB", "#1A5276")
)

plot
```

## 🏷️ Labels and Tooltips

### Basic Tooltips

By default, maps show the region name and value on hover. Add more columns to the tooltip with `tooltip_vars`:

```{r}
state_detailed <- data.frame(
  state = c("CA", "TX", "FL", "NY"),
  value = c(100, 80, 60, 90),
  growth = c(2.1, 3.5, 2.8, 1.2),
  label = c("California", "Texas", "Florida", "New York")
)

plot <- viz_map(
  data = state_detailed,
  join_var = "state",
  value_var = "value",
  map_type = "countries/us/us-all",
  tooltip_vars = c("label", "value", "growth"),
  title = "State Metrics",
  value_label = "Score"
)

plot
```

### Tooltip Parameters

| Parameter | Description | Example |
|-----------|-------------|---------|
| `tooltip_vars` | Columns to include in tooltip | `c("name", "value", "change")` |
| `tooltip_prefix` | Text before main value | `"$"`, `"Score: "` |
| `tooltip_suffix` | Text after main value | `"%"`, `" million"` |
| `value_label` | Label for the legend | `"Population"`, `"Score"` |

```{r}
plot <- viz_map(
  data = state_data,
  join_var = "state",
  value_var = "population_millions",
  map_type = "countries/us/us-all",
  title = "US Population",
  value_label = "Population (millions)",
  tooltip_suffix = " million people",
  color_palette = c("#DEEBF7", "#08519C")
)

plot
```

## 🇪🇺 European Maps

```{r}
european_data <- data.frame(
  country = c("DE", "FR", "GB", "IT", "ES", "PL", "NL", "BE"),
  gdp_per_capita = c(46000, 40000, 42000, 32000, 27000, 15000, 52000, 45000)
)

plot <- viz_map(
  data = european_data,
  join_var = "country",
  value_var = "gdp_per_capita",
  map_type = "custom/europe",
  title = "GDP per Capita in Europe",
  value_label = "GDP per Capita (€)",
  tooltip_prefix = "€",
  color_palette = c("#F7FCB9", "#31A354")
)

plot
```

## 📋 Map Types Reference

| Map Type | Coverage | Join Variable |
|----------|----------|---------------|
| `"custom/world"` | All countries | ISO 3166-1 alpha-3 (USA, GBR, DEU) |
| `"countries/us/us-all"` | US states | State abbreviations (CA, TX, NY) |
| `"custom/europe"` | European countries | ISO 3166-1 alpha-2 (DE, FR, GB) |
| `"countries/us/custom/world"` | World regions | Various codes |

## 🌐 Real-World Examples

### Survey Response Rates by State

```{r}
response_data <- data.frame(
  state = c("CA", "TX", "FL", "NY", "PA", "IL", "OH", "GA", "NC", "MI",
            "NJ", "VA", "WA", "AZ", "MA", "TN", "IN", "MO", "MD", "WI"),
  response_rate = c(45, 38, 42, 51, 44, 47, 39, 35, 41, 43,
                    48, 46, 52, 37, 55, 36, 40, 38, 49, 44)
)

plot <- viz_map(
  data = response_data,
  join_var = "state",
  value_var = "response_rate",
  map_type = "countries/us/us-all",
  title = "Survey Response Rate by State (%)",
  color_palette = c("#FFEDA0", "#BD0026")
)

plot
```

### Global Satisfaction Scores

```{r}
global_survey <- data.frame(
  country = c("USA", "CAN", "MEX", "GBR", "DEU", "FRA", "JPN", "AUS", "BRA", "IND"),
  satisfaction = c(7.2, 7.5, 6.8, 7.1, 6.9, 6.5, 6.2, 7.4, 6.6, 5.8)
)

plot <- viz_map(
  data = global_survey,
  join_var = "country",
  value_var = "satisfaction",
  map_type = "custom/world",
  title = "Customer Satisfaction by Country",
  value_label = "Satisfaction Score",
  tooltip_suffix = " / 10",
  color_palette = c("#FEE8C8", "#E34A33")
)

plot
```

## 📁 Using with create_content()

### Basic Integration

```{r}
content <- create_content(data = state_data, type = "map") %>%
  add_viz(
    join_var = "state",
    value_var = "population_millions",
    map_type = "countries/us/us-all",
    title = "Population Distribution"
  )

content %>% preview()
```

### Multiple Maps with Different Metrics

```{r}
multi_metric <- data.frame(
  state = c("CA", "TX", "FL", "NY", "PA"),
  population = c(39.5, 29.1, 21.5, 20.2, 13.0),
  income = c(75000, 60000, 55000, 70000, 58000),
  satisfaction = c(7.2, 6.8, 7.0, 6.5, 6.9)
)

content <- create_content(data = multi_metric, type = "map", map_type = "countries/us/us-all") %>%
  add_viz(
    join_var = "state",
    value_var = "population",
    title = "Population",
    tabgroup = "Metrics"
  ) %>%
  add_viz(
    join_var = "state",
    value_var = "income",
    title = "Median Income",
    tabgroup = "Metrics"
  ) %>%
  add_viz(
    join_var = "state",
    value_var = "satisfaction",
    title = "Satisfaction",
    tabgroup = "Metrics"
  )
```

## 💡 Data Preparation Tips

### Finding the Right Codes

Maps use standardized geographic codes. Common formats:

| Region Type | Code Format | Examples |
|-------------|-------------|----------|
| Countries (world) | ISO 3166-1 alpha-3 | USA, GBR, DEU, FRA |
| Countries (Europe) | ISO 3166-1 alpha-2 | US, GB, DE, FR |
| US States | USPS abbreviations | CA, TX, NY, FL |

### Handling Missing Regions

Regions without data appear in a neutral color:

```{r}
# Only 5 states - others will be gray
partial_data <- data.frame(
  state = c("CA", "TX", "FL", "NY", "WA"),
  metric = c(100, 90, 85, 95, 80)
)

plot <- viz_map(
  data = partial_data,
  join_var = "state",
  value_var = "metric",
  map_type = "countries/us/us-all",
  title = "Partial Coverage Map",
  value_label = "Score",
  null_color = "#EEEEEE",  # Color for missing regions
  color_palette = c("#C6DBEF", "#084594")
)

plot
```

## 🔍 When to Use Maps

**Use maps when:**
- Showing geographic patterns
- Regional comparisons are important
- Audience thinks spatially about the data
- Location context adds meaning

**Use bar charts instead when:**
- Precise comparisons matter more than geography
- Data isn't inherently geographic
- You have many small regions (hard to see)

## 📚 See Also

- `?viz_map` - Full function documentation
- `vignette("heatmap_vignette")` - For non-geographic grids
- `vignette("content-collections")` - For dashboard integration

## ⚠️ Note on Map Data

Maps require the Highcharts Maps module, which is included with dashboardr. The map geometries are loaded from Highcharts' CDN when the dashboard is viewed. For offline use or custom boundaries, additional configuration may be needed.
