---
title: "Publishing Dashboards to GitHub Pages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Publishing Dashboards to GitHub Pages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette will take you from zero to publishing your dashboard online in under 10 minutes. No prior Git or GitHub knowledge required!

## What You'll Learn

By the end of this guide, you'll be able to:

- Publish your dashboard to GitHub Pages (completely free)
- Update your published dashboard whenever you make changes
- Share your dashboard URL with anyone

## Prerequisites

You need:

- R and RStudio installed
- The `dashboardr` package installed
- A dashboard you want to publish (see the `getting-started` vignette if you need to create one)

```{r setup}
# Install packages if you don't have them
install.packages(c("usethis", "gert"))
library(dashboardr)
```

# Part 1: One-Time Setup (Do This Once)

## Step 1: Install Git on Your Computer

Git is the tool that tracks changes to your files. It needs to be installed before R can use it.

### Windows

1. Download Git from https://git-scm.com/download/win
2. Run the installer (accept all defaults)
3. Restart RStudio after installation

### macOS

Git usually comes pre-installed. To check or update:

1. Open Terminal (find it in Applications > Utilities)
2. Type: `xcode-select --install`
3. Follow the prompts

### Linux

Open terminal and run:

```bash
# Ubuntu/Debian:
sudo apt install git

# Fedora:
sudo dnf install git
```

## Step 2: Tell Git Who You Are

Git needs to know your name and email. Run this in R with **your** information:

```{r configure}
library(usethis)

use_git_config(
  user.name = "Your Name",          # Your actual name
  user.email = "you@example.com"    # Your actual email
)
```

**Important:** Use the same email you'll use for GitHub.

## Step 3: Check Everything Works

Run this command to verify your setup:

```{r check}
git_sitrep()
```

You should see:

- ✓ Your name and email
- ✓ Git version number
- Some other information (warnings are okay for now)

## Step 4: Create a GitHub Account

GitHub will host your dashboard for free.

1. Go to https://github.com
2. Click "Sign up"
3. Follow the steps (choose a username you like - it will be in your dashboard URL!)
4. Verify your email address

## Step 5: Set Up Authentication

GitHub needs a way to verify it's really you. We'll create a Personal Access Token (PAT).

### Create Your Token

Run this command in R:

```{r create-token}
usethis::create_github_token()
```

This will:

1. Open GitHub in your browser
2. Pre-fill a token creation form

On the GitHub page:

1. Add a note like "dashboardr publishing" (to remember what it's for)
2. Click the green "Generate token" button at the bottom
3. **IMPORTANT:** Copy the token that appears (it looks like `ghp_xxxxxxxxxxxx`)
4. Save it somewhere safe - you won't be able to see it again!

### Store Your Token

Now tell R about your token:

```{r store-token}
gitcreds::gitcreds_set()
```

When prompted, paste your token and press Enter.

**That's it!** You've completed the one-time setup. You never have to do these steps again.

# Part 2: Publishing Your Dashboard

Now comes the fun part - publishing your dashboard!

## Step 1: Generate Your Dashboard

First, create and render your dashboard as usual:

```{r generate}
# Example dashboard
my_data <- data.frame(
  category = c("A", "B", "C", "D"),
  value = c(23, 45, 67, 34)
)

viz <- create_viz(type = "bar", x_var = "category") %>%
  add_viz()

dashboard <- create_dashboard(
  output_dir = "my_dashboard",
  title = "My First Dashboard"
) %>%
  add_page(
    name = "Analysis",
    data = my_data,
    visualizations = viz
  )

# Generate the dashboard files
generate_dashboard(dashboard, render = TRUE)
```

## Step 2: Publish!

Now for the magic command:

```{r publish}
publish_dashboard()
```

**What happens:**

1. Git is initialized in your dashboard folder
2. A comprehensive `.gitignore` file is created (automatically excludes data files, large files >10MB, and temporary files)
3. You'll be asked to confirm committing your files (type `1` or `3` when prompted)
4. A GitHub repository is created
5. GitHub Pages is configured
6. Your dashboard is pushed online

**Important notes:**

- You may see prompts asking for confirmation - just type the number to proceed
- Your browser may open showing your GitHub repository
- The function will display your dashboard URL at the end
- GitHub Pages takes **2-5 minutes** to build your site the first time

**Your dashboard URL will be shown in the output and will look like:**
```
https://YOUR-USERNAME.github.io/your-dashboard-name/
```

Copy this URL and wait 2-5 minutes, then visit it to see your live dashboard!

## Publish Options

You can customize the publishing process:

```{r publish-options}
# Create a private repository
publish_dashboard(private = TRUE)

# Publish to an organization
publish_dashboard(organisation = "my-organization")

# Custom initial commit message
publish_dashboard(message = "Launch my awesome dashboard!")

# Use SSH instead of HTTPS
publish_dashboard(protocol = "ssh")
```

# Part 3: Updating Your Dashboard

Made changes to your dashboard? Here's how to update the published version.

## Step 1: Regenerate Your Dashboard

Make your changes and regenerate:

```{r update-generate}
# Make changes to your dashboard
dashboard <- create_dashboard(
  output_dir = "my_dashboard",
  title = "My Updated Dashboard"
) %>%
  add_page(
    name = "Analysis",
    data = my_data,
    visualizations = viz,
    text = "Now with more insights!"  # Added new content
  )

# Regenerate
generate_dashboard(dashboard, render = TRUE)
```

## Step 2: Push Your Updates

```{r update}
update_dashboard()
```

**What happens:**

1. All your changed files are staged
2. Changes are committed with your message
3. **You'll be asked to confirm** before pushing (shows list of files)
4. Changes are pushed to GitHub

**Safety feature:** By default, `update_dashboard()` will show you what files will be pushed and ask for confirmation. Type `yes` or `y` to proceed.

**That's it!** Your changes will be live in 1-2 minutes. The function will display your dashboard URL - just refresh the page to see your updates.

### Update Options

```{r update-options}
# Custom commit message
update_dashboard(message = "Added new visualizations")

# Update specific files only
update_dashboard(
  files = c("docs/index.html", "docs/analysis.html"),
  message = "Updated main pages"
)

# Update all HTML files
update_dashboard(
  files = "docs/*.html",
  message = "Regenerated all pages"
)

# Skip confirmation prompt (not recommended)
update_dashboard(ask = FALSE)
```

# Complete Workflow Example

Here's a full example from start to finish:

```{r complete-workflow}
library(dashboardr)

# 1. Create your dashboard
my_data <- mtcars

viz <- create_viz(type = "bar", x_var = "cyl") %>%
  add_viz(text = "Cars by Cylinder Count")

dashboard <- create_dashboard(
  output_dir = "mtcars_dashboard",
  title = "Motor Trends Car Analysis"
) %>%
  add_page(
    name = "Overview",
    data = my_data,
    visualizations = viz
  )

# 2. Generate the dashboard
generate_dashboard(dashboard, render = TRUE)

# 3. Publish (only needed once!)
publish_dashboard(message = "Initial publication")

# ... time passes, you make changes ...

# 4. Regenerate with changes
generate_dashboard(dashboard, render = TRUE)

# 5. Push updates
update_dashboard(message = "Updated car analysis")
```

# Troubleshooting

## "Error: No GitHub token"

**Problem:** R can't find your GitHub token.

**Solution:** Run `gitcreds::gitcreds_set()` and paste your token again.

## "Error: Git is not installed"

**Problem:** Git isn't installed or R can't find it.

**Solution:** 
1. Install Git (see Step 1)
2. Restart RStudio
3. Run `git_sitrep()` to verify

## "Repository already exists"

**Problem:** You're trying to publish a second time.

**Solution:** Use `update_dashboard()` instead of `publish_dashboard()`.

## "Error: not a git repository"

**Problem:** Git wasn't initialized properly, or you're trying to update before publishing.

**Solution:**
1. Make sure you've run `publish_dashboard()` at least once
2. Run `publish_dashboard()` first to initialize Git and create the repository

## GitHub Pages shows 404

**Problem:** Your site isn't deployed yet, or Pages isn't configured.

**Solution:**
1. Wait 2-5 minutes after first publish
2. Check your repository settings on GitHub
3. Make sure "Pages" is enabled and set to deploy from `/docs` folder

## Data files are being committed

**Problem:** Your data files are being uploaded to GitHub.

**Solution:** The `.gitignore` file should automatically exclude data files. Check:
1. Look for a `.gitignore` file in your dashboard folder
2. Make sure your data files match the patterns (*.csv, *.rds, etc.)
3. Large files (>10MB) are automatically detected and ignored

# Tips and Best Practices

## 1. Keep Data Out of Git

The publishing functions automatically exclude:
- All common data file formats (CSV, RDS, Excel, etc.)
- Large files (>10MB)
- Data directories

This keeps your repository small and protects sensitive data.

## 2. Write Descriptive Commit Messages

Instead of:
```{r bad-message}
update_dashboard(message = "update")
```

Do this:
```{r good-message}
update_dashboard(message = "Added new bar chart for regional analysis")
```

## 3. Test Locally First

Always run `generate_dashboard(dashboard, render = TRUE, open = "browser")` to preview your changes before publishing.

## 4. Update Regularly

It's easier to push small, frequent updates than large, infrequent ones. Update after each significant change.

## 5. Check Your Site

After updating, wait 1-2 minutes then visit your dashboard URL to verify the changes appear correctly.

# Next Steps

Now that your dashboard is published, you can:

- Share the URL with colleagues, clients, or the public
- Add more pages and visualizations
- Customize your dashboard theme
- Set up custom domains (see GitHub Pages documentation)

**Happy publishing!** 🎉

For more advanced features, see:
- `vignette("advanced-features", package = "dashboardr")`
- GitHub Pages documentation: https://docs.github.com/pages

