---
title: "Creating Scatter Plots with viz_scatter()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Scatter Plots with viz_scatter()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## 📖 Introduction

The `viz_scatter()` function visualizes the relationship between two numeric variables. Each point represents one observation, positioned by its x and y values. Essential for exploring correlations, clusters, and outliers.

```{r setup}
library(dashboardr)
library(dplyr)
library(gssr)
library(haven)

# Load GSS data - we need numeric variables for scatter plots
data(gss_all)
gss <- gss_all %>%
  select(year, age, sex, race, degree, happy, polviews, educ, realinc) %>%
  filter(year == 2022,  # Use 2022 which has realinc data
         !is.na(age), !is.na(educ), !is.na(realinc),
         realinc > 0, educ > 0) %>%
  mutate(
    sex = droplevels(as_factor(sex)),
    degree = droplevels(as_factor(degree))
  )
```

```{r, include=FALSE}
devtools::load_all()
```

## 📊 Basic Scatter Plots

Create a simple scatter plot showing the relationship between education (years) and income:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "educ",
  y_var = "realinc",
  title = "Education vs Income",
  x_label = "Years of Education",
  y_label = "Real Income ($)"
)

plot
```

## 📈 Adding Trend Lines

Use `show_trend = TRUE` to add a regression line:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "educ",
  y_var = "realinc",
  show_trend = TRUE,
  title = "Education vs Income (with trend)"
)

plot
```

## 🎨 Coloring by Groups

Use `color_var` to color points by a categorical variable:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "educ",
  y_var = "realinc",
  color_var = "sex",
  title = "Education vs Income by Sex",
  color_palette = c("#3498DB", "#E74C3C")
)

plot
```

## 📊 Age vs Education

Another relationship to explore - age and years of education:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "age",
  y_var = "educ",
  color_var = "degree",
  title = "Age vs Education by Degree",
  x_label = "Age (years)",
  y_label = "Years of Education",
  alpha = 0.5,
  color_palette = c("#E74C3C", "#F39C12", "#27AE60", "#3498DB", "#9B59B6")
)

plot
```

## 👁️ Handling Overlap with Transparency

For dense data, use `alpha` to reveal patterns:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "age",
  y_var = "realinc",
  alpha = 0.3,
  point_size = 3,
  title = "Age vs Income (with transparency)"
)

plot
```

## 🏷️ Labels and Tooltips

Customize axis labels and tooltip formatting for better readability:

```{r}
plot <- viz_scatter(
  data = gss,
  x_var = "educ",
  y_var = "realinc",
  title = "Education vs Income",
  x_label = "Years of Education",
  y_label = "Annual Income (USD)",
  tooltip_format = "Education: {x} years, Income: ${y}"
)

plot
```

| Parameter | Description | Example |
|-----------|-------------|---------|
| `x_label` | Custom x-axis label | `"Years of Education"` |
| `y_label` | Custom y-axis label | `"Income (USD)"` |
| `tooltip_format` | Custom tooltip template | `"x: {x}, y: {y}"` |

The `tooltip_format` parameter supports placeholders: `{x}` for x-value, `{y}` for y-value, and `{color}` for the color group.

## 📁 Using with create_content()

Integrate scatter plots into dashboards:

```{r}
content <- create_content(data = gss, type = "scatter") %>%
  add_viz(
    x_var = "educ",
    y_var = "realinc",
    show_trend = TRUE,
    title = "Education vs Income"
  )

content %>% preview()
```

### With Filters

Compare relationships across groups:

```{r}
content <- create_content(data = gss, type = "scatter", alpha = 0.5) %>%
  add_viz(
    x_var = "educ",
    y_var = "realinc",
    title = "Male",
    filter = ~ sex == "male",
    tabgroup = "By Sex"
  ) %>%
  add_viz(
    x_var = "educ",
    y_var = "realinc",
    title = "Female",
    filter = ~ sex == "female",
    tabgroup = "By Sex"
  )

content %>% preview()
```

### Multiple Relationships

```{r}
content <- create_content(data = gss, type = "scatter", alpha = 0.4, show_trend = TRUE) %>%
  add_viz(
    x_var = "educ",
    y_var = "realinc",
    title = "Education → Income",
    tabgroup = "Relationships"
  ) %>%
  add_viz(
    x_var = "age",
    y_var = "realinc",
    title = "Age → Income",
    tabgroup = "Relationships"
  )

content %>% preview()
```

## 🔍 Interpreting Scatter Plots

### Correlation Patterns

| Pattern | Meaning |
|---------|---------|
| Upward slope | Positive relationship |
| Downward slope | Negative relationship |
| No pattern | No linear relationship |
| Tight cluster | Strong relationship |
| Wide scatter | Weak relationship |

### What to Look For

1. **Direction** - Positive, negative, or none?
2. **Strength** - How tightly clustered?
3. **Linearity** - Linear or curved?
4. **Outliers** - Points far from the pattern?
5. **Clusters** - Distinct groups?

## 💡 When to Use Scatter Plots

**Use `viz_scatter()` when:**
- Exploring relationship between two numeric variables
- Looking for correlations
- Identifying outliers
- Showing individual-level data

**Use `viz_histogram()` when:**
- Showing distribution of a single variable

**Use `viz_heatmap()` when:**
- Data is aggregated (means, counts)
- Many overlapping points

## 📚 See Also

- `?viz_scatter` - Full function documentation
- `vignette("histogram_vignette")` - For single-variable distributions
- `vignette("heatmap_vignette")` - For aggregated relationships
- `vignette("content-collections")` - For dashboard integration
