---
title: "Showcase Dashboard Code Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Showcase Dashboard Code Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  purl = FALSE
)
```

<div style="background-color: #e7f3fe; border-left: 6px solid #2196F3; padding: 15px; margin-bottom: 20px;">
<h3 style="margin-top: 0;">🚀 View the Live Demo</h3>
<p>See this code in action! <strong><a href="../live-demos/showcase/index.html">Click here to view the Showcase Dashboard live demo →</a></strong></p>
<p>Or run it yourself: <code>dashboardr::showcase_dashboard()</code></p>
</div>

## Overview

This vignette provides the **exact R code** for each visualization in the Showcase Dashboard. Each section shows the complete `add_viz()` call as it appears in the `showcase_dashboard()` function source code.

Click on any visualization in the live demo to see the "View R Code" accordion, which links back to the corresponding section here.

---

## Data Preparation {#data-prep}

```{r data-prep-code, purl = FALSE}
library(dashboardr)
library(dplyr)

# Load GSS panel data
data(gss_panel20, package = "gssr")

# Clean and select relevant variables
gss_clean <- gss_panel20 %>%
  select(
    age_1a, sex_1a, degree_1a, region_1a,
    happy_1a, trust_1a, fair_1a, helpful_1a,
    polviews_1a, partyid_1a, class_1a
  ) %>%
  filter(if_any(everything(), ~ !is.na(.)))
```

---

## Demographics & Education Tabset

### Happiness by Education {#demographics-happiness-education}

```{r demographics-happiness-education-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "degree_1a",
        stack_var = "happy_1a",
        title = "Happiness Distribution Across Education Levels",
        subtitle = "Percentage breakdown within each education category",
        x_label = "Education Level",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "How happy are you with your life right now?",
        text_position = "above",
        icon = "ph:chart-bar",
        height = 500,
        tabgroup = "demographics")
```

---

### Happiness by Gender {#demographics-happiness-gender}

```{r demographics-happiness-gender-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "sex_1a",
        stack_var = "happy_1a",
        title = "Happiness Distribution by Gender",
        subtitle = "Gender differences in reported happiness levels",
        x_label = "Gender",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "How happy are you with your life right now?",
        text_position = "below",
        icon = "ph:gender-intersex",
        height = 450,
        tabgroup = "demographics")
```

---

## Political Attitudes Tabset

### Party ID by Political Views {#politics-party-views}

```{r politics-party-views-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "polviews_1a",
        stack_var = "partyid_1a",
        title = "Party ID by Political Views",
        subtitle = "Distribution of party identification across political ideology",
        x_label = "Political Views",
        y_label = "Percentage of Respondents",
        stack_label = "Party ID",
        stacked_type = "percent",
        tooltip_suffix = "%",
        color_palette = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2"),
        text = "This shows how party identification aligns with political ideology.",
        text_position = "above",
        icon = "ph:users-three",
        height = 550,
        tabgroup = "politics")
```

---

### Trust by Region {#politics-trust-region}

```{r politics-trust-region-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "region_1a",
        stack_var = "trust_1a",
        title = "Trust Levels by US Region",
        subtitle = "Regional variation in interpersonal trust",
        x_label = "US Region",
        y_label = "Percentage of Respondents",
        stack_label = "Trust Level",
        stack_order = c("Can Trust", "Can't Be Too Careful", "Depends"),
        stacked_type = "percent",
        tooltip_suffix = "%",
        color_palette = c("#2E8B57", "#DAA520", "#CD5C5C"),
        text = "Do you think you can usually trust strangers?",
        text_position = "below",
        icon = "ph:map-pin",
        height = 500,
        tabgroup = "politics")
```

---

### Gender by Social Class {#politics-gender-class}

```{r politics-gender-class-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "class_1a",
        stack_var = "sex_1a",
        title = "Gender Distribution Across Social Classes",
        subtitle = "With custom labels and ordering",
        x_label = "Self-Reported Social Class",
        y_label = "Number of Respondents",
        stack_label = "Gender",
        x_order = c("Lower Class", "Working Class", "Middle Class", "Upper Class"),
        stack_order = c("Female", "Male"),
        stacked_type = "counts",
        tooltip_prefix = "Count: ",
        color_palette = c("#E07A5F", "#3D5A80"),
        text = "Gender distribution across different social class categories.",
        text_position = "above",
        icon = "ph:chart-pie",
        height = 450,
        tabgroup = "politics")
```

**Note:** This chart uses `stacked_type = "counts"` instead of "percent".

---

## Social Issues Tabset

### Trust by Education and Age {#social-trust-education-age}

```{r social-trust-education-age-code, purl = FALSE}
add_viz(type = "heatmap",
        x_var = "degree_1a",
        y_var = "age_1a",
        value_var = "trust_1a",
        title = "Trust by Education and Age",
        subtitle = "Average trust levels across education and age groups",
        x_label = "Education Level",
        y_label = "Age Group",
        value_label = "Trust Level",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "This heatmap reveals trust patterns across education and age groups.",
        text_position = "below",
        icon = "ph:heatmap",
        height = 600,
        tabgroup = "social")
```

---

### Trust by Region and Education {#social-trust-region-education}

```{r social-trust-region-education-code, purl = FALSE}
add_viz(type = "heatmap",
        x_var = "region_1a",
        y_var = "degree_1a",
        value_var = "trust_1a",
        title = "Trust by Region and Education",
        subtitle = "Educational and regional patterns in trust levels",
        x_label = "Region",
        y_label = "Education Level",
        value_label = "Trust Level",
        y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text = "Educational and regional patterns in trust distribution.",
        text_position = "above",
        icon = "ph:chart-pie",
        height = 550,
        tabgroup = "social")
```

---

## Standalone Summary Charts

### Overall Happiness by Education {#summary-happiness-education}

```{r summary-happiness-education-code, purl = FALSE}
add_viz(type = "stackedbar",
        x_var = "degree_1a",
        stack_var = "happy_1a",
        title = "Overall Happiness by Education",
        subtitle = "Complete distribution of happiness across education levels",
        x_label = "Education Level",
        y_label = "Percentage of Respondents",
        stack_label = "Happiness Level",
        stacked_type = "percent",
        x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
        stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
        tooltip_suffix = "%",
        color_palette = c("#2E86AB", "#A23B72", "#F18F01"),
        text = "How happy are you?",
        text_position = "above",
        icon = "ph:chart-bar",
        height = 600)
```

---

### Overall Trust by Politics {#summary-trust-politics}

```{r summary-trust-politics-code, purl = FALSE}
add_viz(type = "heatmap",
        x_var = "partyid_1a",
        y_var = "polviews_1a",
        value_var = "trust_1a",
        title = "Overall Trust by Politics",
        subtitle = "Complete trust patterns across political groups",
        x_label = "Party Identification",
        y_label = "Political Views",
        value_label = "Trust Level",
        x_order = c("Strong Democrat", "Not Very Strong Democrat", "Independent, Close to Democrat",
                    "Independent", "Independent, Close to Republican", "Not Very Strong Republican", "Strong Republican"),
        y_order = c("Extremely Liberal", "Liberal", "Slightly Liberal", "Moderate",
                    "Slightly Conservative", "Conservative", "Extremely Conservative"),
        color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
        tooltip_prefix = "Trust: ",
        tooltip_suffix = "/3",
        tooltip_labels_format = "{point.value:.2f}",
        text_position = "below",
        icon = "ph:shield-check",
        height = 700)
```

---

## Visualization Index

| Chart | Type | Page | Tabset | Anchor |
|-------|------|------|--------|--------|
| Happiness by Education | Stacked Bar | GSS Data Analysis | Demographics | [#demographics-happiness-education](#demographics-happiness-education) |
| Happiness by Gender | Stacked Bar | GSS Data Analysis | Demographics | [#demographics-happiness-gender](#demographics-happiness-gender) |
| Party ID by Political Views | Stacked Bar | GSS Data Analysis | Politics | [#politics-party-views](#politics-party-views) |
| Trust by Region | Stacked Bar | GSS Data Analysis | Politics | [#politics-trust-region](#politics-trust-region) |
| Gender by Social Class | Stacked Bar | GSS Data Analysis | Politics | [#politics-gender-class](#politics-gender-class) |
| Trust by Education & Age | Heatmap | GSS Data Analysis | Social Issues | [#social-trust-education-age](#social-trust-education-age) |
| Trust by Region & Education | Heatmap | GSS Data Analysis | Social Issues | [#social-trust-region-education](#social-trust-region-education) |
| Overall Happiness | Stacked Bar | Summary Charts | - | [#summary-happiness-education](#summary-happiness-education) |
| Overall Trust by Politics | Heatmap | Summary Charts | - | [#summary-trust-politics](#summary-trust-politics) |

## Next Steps

- **[View the live Showcase Dashboard →](../live-demos/showcase/index.html)**
- See the [Tutorial Dashboard Code](tutorial_dashboard_code.html) for simpler examples
- Explore [visualization vignettes](../articles/index.html) for detailed chart customization
