---
title: "Getting Started With `viz_timeline()`"
author: "Alexandra Pafford"
date: "2025-08-14"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating interactive timeline visualizations with `viz_timeline`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, # Default figure width
  fig.height = 5, # Default figure height
  warning = FALSE, # Suppress warnings from highcharter, haven etc. for cleaner output
  message = FALSE  # Suppress messages (e.g., haven conversion notes) for cleaner output
)
```

## Introduction

The `viz_timeline()` function creates interactive timeline visualizations for survey data, particularly useful for showing changes in Likert-type responses over time. Furthermore, the function has been designed to handle SPSS (.sav) data as well, which makes it very handy for researchers who are accustomed to working with social science data in this format.

This vignette demonstrates how to use the function with the General Social Survey (GSS) data. Because we are working with data over time, we will use the `gss_all` data set. This is a large data set, so it might take a while to load.

## Setup

First, let's load the required packages and define the function:

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(highcharter)
library(stringr)
library(haven)
library(dashboardr)
library(gssr)

# Load GSS data
data(gss_all)
```

## Example 1: Basic Stacked Area Chart

Let's start with a simple stacked area chart showing confidence in financial institutions over time:

```{r example1}
plot1 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "confinan",
  chart_type = "stacked_area",
  title = "Confidence in Financial Institutions Over Time",
  y_max = 100
)

plot1
```

This chart shows how public confidence in financial institutions has changed from the 1970s to recent years, with each colored area representing a different level of confidence.

## Example 2: Line Chart with Grouping

Now let's create a line chart showing happiness trends by gender:

```{r example2}
plot2 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "happy",
  group_var = "sex",
  chart_type = "line",
  title = "Happiness Trends by Gender",
  y_levels = c("very happy", "pretty happy", "not too happy")
)

plot2
```

This line chart displays separate lines for each combination of happiness level and gender, allowing us to compare trends between men and women over time.

## Example 3: Time Binning

For data spanning many years, we can bin the time variable into decades:

```{r example3}
plot3 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "satfin",
  chart_type = "stacked_area",
  title = "Financial Satisfaction by Decade",
  time_breaks = c(1970, 1980, 1990, 2000, 2010, 2020),
  time_bin_labels = c("1970s", "1980s", "1990s", "2000s", "2010s"),
  y_max = 100
)

plot3
```

This approach is useful when you want to show broader trends across time periods rather than year-by-year changes.

## Example 4: Controlling Response Order

You can control the order of response categories to ensure logical ordering:

```{r example4}
plot4 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "health",
  chart_type = "stacked_area",
  title = "Self-Reported Health Over Time",
  y_levels = c("poor", "fair", "good", "excellent"),
  y_max = 100
)

plot4
```

By specifying `y_levels`, we ensure that health categories are ordered from worst to best, making the chart more intuitive to read.

## Tips for Using the Function

### 1. Check Your Data First

Before creating charts, it's helpful to examine your data:

```{r data-check}
# Check available variables
names(gss_all)[1:20]

# Check response levels for a variable
gss_all %>%
  select(happy) %>%
  filter(!is.na(happy)) %>%
  mutate(happy = haven::as_factor(happy, levels = "labels")) %>%
  count(happy)
```

### 2. Handle Missing Data

The function automatically filters out missing values, but you should be aware of how much data is being excluded:

```{r missing-data}
# Check data availability
gss_all %>%
  summarise(
    total_rows = n(),
    year_missing = sum(is.na(year)),
    happy_missing = sum(is.na(happy)),
    both_available = sum(!is.na(year) & !is.na(happy))
  )
```

### 3. Interactive Features

The resulting charts are interactive Highcharts objects that support:

- **Hovering** over data points to see exact values
- **Clicking legend items** to show/hide series
- **Zooming and panning** for detailed exploration
- **Exporting** charts as images or data

### 4. Labels and Tooltips

Customize axis labels for publication-ready charts:

```{r labels-example}
plot_labeled <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "happy",
  chart_type = "stacked_area",
  title = "Happiness Trends in America",
  x_label = "Survey Year",
  y_label = "Percentage of Respondents",
  y_max = 100
)

plot_labeled
```

| Parameter | Description | Default |
|-----------|-------------|---------|
| `x_label` | Custom x-axis label | Variable name (`time_var`) |
| `y_label` | Custom y-axis label | `"Percentage"` |

The tooltips in timeline charts automatically show the year, category, and percentage when hovering over data points.

## Advanced Features

### Response Binning

Bin numeric responses into categories for clearer visualization:

```{r response-binning}
# Bin responses into positive/neutral/negative
plot5 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "satfin",  # Financial satisfaction (1-3 scale)
  chart_type = "stacked_area",
  title = "Financial Satisfaction (Binned)",
  y_breaks = c(0.5, 1.5, 2.5, 3.5),
  y_bin_labels = c("Not satisfied", "More or less", "Satisfied"),
  y_max = 100
)

plot5
```

**When to use:**
- Simplify complex response scales
- Group similar responses together
- Create more interpretable categories

### Response Filtering

Focus on specific response values:

```{r response-filtering}
# Show only "very happy" and "pretty happy" responses
plot6 <- viz_timeline(
  data = gss_all,
  time_var = "year",
  y_var = "happy",
  chart_type = "line",
  title = "Positive Happiness Trends",
  y_filter = c("very happy", "pretty happy")
)

plot6
```

For numeric scales, use range notation:

```{r response-filter-numeric, eval=FALSE}
# Show only top-box scores (5-7 on 1-7 scale)
viz_timeline(
  data = survey_data,
  time_var = "wave",
  y_var = "satisfaction",
  y_filter = 5:7,
  y_filter_combine = TRUE,
  y_filter_label = "Top 3 Box (5-7)"
)
```

**Parameters:**
- `y_filter`: Which values to include
- `y_filter_combine`: Combine filtered values into single line (default: TRUE)
- `y_filter_label`: Custom label when combined (default: auto-generated)

**Calculation Note:** When `y_filter_combine = TRUE`, percentages are calculated out of the *total* responses (not just filtered ones), showing the true proportion of top-box scores.

### Combining Filters with Groups

Show filtered responses across multiple groups:

```{r filter-with-groups, eval=FALSE}
# Top-box satisfaction by age group
viz_timeline(
  data = survey_data,
  time_var = "year",
  y_var = "satisfaction",
  group_var = "age_group",
  y_filter = 5:7,
  y_filter_combine = TRUE,
  y_filter_label = "Highly Satisfied",
  title = "High Satisfaction by Age Group"
)
```

This creates separate lines for each age group, all showing only the highly satisfied responses.

## Integration with Dashboards

The `viz_timeline()` function works seamlessly with `dashboardr`:

```{r dashboard-integration, eval=FALSE}
library(dashboardr)

# Create multiple timeline visualizations
timeline_viz <- create_viz() %>%
  add_viz(
    type = "timeline",
    time_var = "year",
    y_var = "happy",
    chart_type = "line",
    title = "Happiness Trends",
    tabgroup = "trends"
  ) %>%
  add_viz(
    type = "timeline",
    time_var = "year",
    y_var = "satfin",
    chart_type = "stacked_area",
    title = "Financial Satisfaction",
    tabgroup = "trends"
  )

# Create dashboard
dashboard <- create_dashboard(
  title = "GSS Trends Dashboard",
  output_dir = "gss_dashboard"
) %>%
  add_page(
    "Trends",
    data = gss_all,
    visualizations = timeline_viz,
    is_landing_page = TRUE
  )

generate_dashboard(dashboard)
```

### Comparative Analysis with Filters

Use dashboard filters to compare different time periods or demographics:

```{r dashboard-filters, eval=FALSE}
# Compare trends across different demographics
timeline_viz <- create_viz(
  type = "timeline",
  time_var = "year",
  y_var = "happy",
  chart_type = "line"
) %>%
  add_viz(title = "All Respondents") %>%
  add_viz(title = "Men", filter = ~ sex == "male") %>%
  add_viz(title = "Women", filter = ~ sex == "female") %>%
  add_viz(title = "Young Adults", filter = ~ age >= 18 & age <= 35)

dashboard <- create_dashboard(...) %>%
  add_page("Happiness Analysis", 
           data = gss_all, 
           visualizations = timeline_viz)

generate_dashboard(dashboard)
```

This creates separate tabs for each filtered view, making it easy to compare trends across groups.

## Conclusion

The `viz_timeline()` function provides a flexible way to visualize survey data trends over time. The function handles the data processing and creates interactive visualizations that are perfect for exploring temporal patterns in survey responses.

The function is especially well-suited for:

- **Longitudinal survey analysis**
- **Public opinion research**
- **Social trend visualization**
- **Comparative analysis across groups**
- **Interactive reporting and dashboards**

This vignette provides a comprehensive guide to using your function with real GSS data, including practical examples and tips for effective usage.
