---
title: "Getting Started With `create_timeline()`"
author: "Alexandra Pafford"
date: "14/08/2025"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette:
  %\VignetteIndexEntry{Getting Started With `create_timeline()`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8"}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, # Default figure width
  fig.height = 5, # Default figure height
  warning = FALSE, # Suppress warnings from highcharter, haven etc. for cleaner output
  message = FALSE  # Suppress messages (e.g., haven conversion notes) for cleaner output
)
```

## Introduction

The `create_timeline()` function creates interactive timeline visualizations for survey data, particularly useful for showing changes in Likert-type responses over time. Furthermore, the function has been designed to handle SPSS (.sav) data as well, which makes it very handy for researchers who are accustomed to working with social science data in this format.

This vignette demonstrates how to use the function with the General Social Survey (GSS) data. Because we are working with data over time, we will use the `gss_all` data set. This is a large data set, so it might take a while to load.

## Setup

First, let's load the required packages and define the function:

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(highcharter)
library(stringr)
library(haven)
#library(dashboardr)

# Define the %||% operator
`%||%` <- function(x, y) if (is.null(x)) y else x

# Load GSS data
data(gss_all)
```

## Example 1: Basic Stacked Area Chart

Let's start with a simple stacked area chart showing confidence in financial institutions over time:

```{r example1}
plot1 <- create_timeline(
  data = gss_all,
  time_var = "year",
  response_var = "confinan",
  chart_type = "stacked_area",
  title = "Confidence in Financial Institutions Over Time",
  y_max = 100
)

plot1
```

This chart shows how public confidence in financial institutions has changed from the 1970s to recent years, with each colored area representing a different level of confidence.

## Example 2: Line Chart with Grouping

Now let's create a line chart showing happiness trends by gender:

```{r example2}
plot2 <- create_timeline(
  data = gss_all,
  time_var = "year",
  response_var = "happy",
  group_var = "sex",
  chart_type = "line",
  title = "Happiness Trends by Gender",
  response_levels = c("very happy", "pretty happy", "not too happy")
)

plot2
```

This line chart displays separate lines for each combination of happiness level and gender, allowing us to compare trends between men and women over time.

## Example 3: Time Binning

For data spanning many years, we can bin the time variable into decades:

```{r example3}
plot3 <- create_timeline(
  data = gss_all,
  time_var = "year",
  response_var = "satfin",
  chart_type = "stacked_area",
  title = "Financial Satisfaction by Decade",
  time_breaks = c(1970, 1980, 1990, 2000, 2010, 2020),
  time_bin_labels = c("1970s", "1980s", "1990s", "2000s", "2010s"),
  y_max = 100
)

plot3
```

This approach is useful when you want to show broader trends across time periods rather than year-by-year changes.

## Example 4: Controlling Response Order

You can control the order of response categories to ensure logical ordering:

```{r example4}
plot4 <- create_timeline(
  data = gss_all,
  time_var = "year",
  response_var = "health",
  chart_type = "stacked_area",
  title = "Self-Reported Health Over Time",
  response_levels = c("poor", "fair", "good", "excellent"),
  y_max = 100
)

plot4
```

By specifying `response_levels`, we ensure that health categories are ordered from worst to best, making the chart more intuitive to read.

## Tips for Using the Function

### 1. Check Your Data First

Before creating charts, it's helpful to examine your data:

```{r data-check}
# Check available variables
names(gss_all)[1:20]

# Check response levels for a variable
gss_all %>%
  select(happy) %>%
  filter(!is.na(happy)) %>%
  mutate(happy = haven::as_factor(happy, levels = "labels")) %>%
  count(happy)
```

### 2. Handle Missing Data

The function automatically filters out missing values, but you should be aware of how much data is being excluded:

```{r missing-data}
# Check data availability
gss_all %>%
  summarise(
    total_rows = n(),
    year_missing = sum(is.na(year)),
    happy_missing = sum(is.na(happy)),
    both_available = sum(!is.na(year) & !is.na(happy))
  )
```

### 3. Interactive Features

The resulting charts are interactive Highcharts objects that support:

- **Hovering** over data points to see exact values
- **Clicking legend items** to show/hide series
- **Zooming and panning** for detailed exploration
- **Exporting** charts as images or data

## Conclusion

The `create_timeline()` function provides a flexible way to visualize survey data trends over time. The function handles the data processing and creates interactive visualizations that are perfect for exploring temporal patterns in survey responses.

The function is especially well-suited for:

- **Longitudinal survey analysis**
- **Public opinion research**
- **Social trend visualization**
- **Comparative analysis across groups**
- **Interactive reporting and dashboards**

This vignette provides a comprehensive guide to using your function with real GSS data, including practical examples and tips for effective usage.
