---
title: "Welcome to Dashboardr"
format: html
editor: visual
---

```{r setup}
#| include: false
#library(dashboardr)
library(highcharter)
library(dplyr)
library(gssr)

# Load the data
data <- readRDS("dashboard_data.rds")
# Or if using SPSS data
spss_data <-  read.spss("dashboard_data.sav")
```

# Welcome to Dashboardr!

Dashboardr enables you to visualize data in a clean and straightforward way. It makes it easy to create beautiful plots, especially when you have little experience with R. To learn more about Dashboardr, check out our [GitHub](https://github.com/favstats/dashboardr).

This is an example template of what your Dashboardr can look like. When you render, an HTML webpage with your data dashboard will be generated.

# Visualizing Your Data

The plots are easily generated when you use the visualization functions from the `dashboardr` package. Here are a few examples using the `gss_panel20` dataset.

# Educational Attainment by Gender

```{r stackedbar}
# Create basic stacked bar chart
plot1 <- create_stackedbar(
  data = gss_clean,
  x_var = "degree_1a",
  stack_var = "sex_1a",
  title = "Educational Attainment by Gender",
  subtitle = "GSS Panel 2016 - Raw counts",
  x_label = "Highest Degree Completed",
  y_label = "Number of Respondents",
  stack_label = "Gender",
  stacked_type = "normal"
)

plot1
```


# Social Attitudes and Trust

```{r stackedbars}
# Define our questions and labels
social_questions <- c("trust_1a", "fair_1a", "helpful_1a")
social_labels <- c(
  "Interpersonal Trust",
  "Fairness of Others", 
  "Helpfulness of Others"
)

# Create basic multi-question chart
plot2 <- create_stackedbars(
  data = gss_likert,
  questions = social_questions,
  question_labels = social_labels,
  title = "Social Attitudes and Trust",
  subtitle = "GSS Panel 2016 - Distribution of responses across social attitude questions",
  x_label = "Social Attitude Dimension",
  stack_label = "Response Level",
  stacked_type = "normal"
)

plot2
```
