---
title: "{{< iconify ph info >}} About"
format: html
---

```{r}
#| include: false
library(dashboardr)
```

```{=html}
<style>
/* Ensure chart containers expand to fit content - aggressive overrides */
.cell-output-display,
.cell-output,
.panel-tabset-tabby > .tab-content,
.panel-tabset > .tab-content,
.tab-pane,
.tab-pane.active,
.card-body,
.quarto-figure,
section {
  overflow: visible !important;
  height: auto !important;
  max-height: none !important;
}
/* Ensure highcharts containers respect their set height */
.highcharts-container,
.html-widget,
.htmlwidget {
  overflow: visible !important;
}
</style>
<script>
// Aggressive reflow for Highcharts - multiple attempts to catch widget initialization
(function() {
  function reflowCharts() {
    if (typeof Highcharts !== 'undefined' && Highcharts.charts) {
      Highcharts.charts.forEach(function(chart) {
        if (chart) {
          try { chart.reflow(); } catch(e) {}
        }
      });
    }
  }
  
  // Try multiple times at different intervals
  [0, 100, 250, 500, 1000, 2000].forEach(function(delay) {
    setTimeout(function() {
      window.dispatchEvent(new Event('resize'));
      requestAnimationFrame(reflowCharts);
    }, delay);
  });
  
  // Also reflow on any tab click
  document.addEventListener('click', function(e) {
    if (e.target.matches('.nav-link, [data-bs-toggle="tab"]')) {
      setTimeout(reflowCharts, 50);
    }
  });
})();
</script>
```

This dashboard aggregates and visualizes data collected via the General Social Survey (GSS) - 
a nationally representative survey of adults in the United States conducted since 1972.
The data is open-source and you can find out more about the GSS here.

## Variables Used

- **Happiness**: Self-reported happiness levels
- **Trust**: General social trust measures
- **Education**: Educational attainment levels
- **Political Views**: Liberal-conservative scale
- **Party ID**: Political party identification
- **Demographics**: Age, gender, region

## Dashboard Creators

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(htmltools)
library(dashboardr)

mario_card <- card(
  content = "Mario il Gatto is a data scientist who believes that every dataset has a soul and that R is the language of the gods. He spends his days making beautiful visualizations and his nights dreaming of perfectly normalized databases.",
  title = "Mario il Gatto",
  image = "https://images.unsplash.com/photo-1514888286974-6c03e2ca1dba?w=300&h=200&fit=crop",
  image_alt = "Photo of a majestic orange cat",
  footer = "Website: mario-il-gatto.data",
  class = "mb-3"
)

giuseppe_card <- card(
  content = "Giuseppe il Cane is a machine learning engineer who thinks that neural networks are just very complicated dogs. He's convinced that every algorithm needs a good walk and that overfitting is just a sign of too much enthusiasm.",
  title = "Giuseppe il Cane",
  image = "https://images.unsplash.com/photo-1552053831-71594a27632d?w=300&h=200&fit=crop",
  image_alt = "Photo of a happy golden retriever",
  footer = "Website: giuseppe-il-cane.ai",
  class = "mb-3"
)

# Display cards in a row using the card_row function
card_row(mario_card, giuseppe_card)
```
## More about Dashboardr
Dashboardr is an R package with a clear vision: to make it intuitive for everyone to create beautiful dashboards.
The package is especially useful when time is limited. The iterative piping logic means that it is very quick to add new pages
and plots, even when the user is inexperienced with programming.
In a variety of contexts, this means that you can get quick, beautiful insights to present findings to wider audiences.

