---
title: "{{< iconify ph chart-line >}} Example Dashboard"
format: html
---

Here, you can see how to add text within a dashboard.

## Add a new heading like this

A line break is displayed when you add a new section.

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load required libraries
library(dashboardr)
library(dplyr)
library(highcharter)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

# Load data from dataset_2867obs.rds
data <- readRDS('dataset_2867obs.rds')

# Data summary
cat('Dataset loaded:', nrow(data), 'rows,', ncol(data), 'columns\n')

```

## Example 1: Stacked Bars


::: {.panel-tabset}

### {{< iconify ph chart-bar >}} Change the title here...


If you want to add text within the tab, you can do so here.

```{r demographics}
# Change the title here...
result <- create_stackedbar(
  data = data,
  title = "Change the title here...",
  x_var = "degree_1a",
  stack_var = "happy_1a",
  subtitle = "You can add a subtitle using the subtitle argument",
  x_label = "Education Level",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Apply height to highcharter object
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_chart(result, height = 500)
}

result
```


### {{< iconify ph gender-intersex >}} Tabset #2

```{r demographics-2}
# Tabset #2
result <- create_stackedbar(
  data = data,
  title = "Tabset #2",
  x_var = "sex_1a",
  stack_var = "happy_1a",
  subtitle = "Another example subtitle here!",
  x_label = "Gender",
  y_label = "Percentage of Respondents",
  stack_label = "Happiness Level",
  stacked_type = "percent",
  stack_order = c("Very Happy", "Pretty Happy", "Not Too Happy"),
  tooltip_suffix = "%",
  color_palette = c("#2E86AB", "#A23B72", "#F18F01")
)

# Apply height to highcharter object
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_chart(result, height = 450)
}

result
```

Change the position of the text using the `text_position` argument.



:::

## Example 2: Heatmap


::: {.panel-tabset}

### {{< iconify ph heatmap >}} Trust by Education and Age


Here's another example of the kind of plots you can generate in your dashboard.

```{r social}
# Trust by Education and Age
result <- create_heatmap(
  data = data,
  title = "Trust by Education and Age",
  x_var = "degree_1a",
  y_var = "age_1a",
  value_var = "trust_1a",
  subtitle = "Average trust levels across education and age groups",
  x_label = "Example x axis",
  y_label = "Customizable y label",
  value_label = "You can change the label here too...",
  x_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Apply height to highcharter object
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_chart(result, height = 600)
}

result
```


### {{< iconify ph chart-pie >}} Trust by Region and Education


Educational and regional patterns in trust distribution.

```{r social-2}
# Trust by Region and Education
result <- create_heatmap(
  data = data,
  title = "Trust by Region and Education",
  x_var = "region_1a",
  y_var = "degree_1a",
  value_var = "trust_1a",
  subtitle = "Educational and regional patterns in trust levels",
  x_label = "Region",
  y_label = "Education Level",
  value_label = "Trust Level",
  y_order = c("Lt High School", "High School", "Junior College", "Bachelor", "Graduate"),
  color_palette = c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"),
  tooltip_prefix = "Trust: ",
  tooltip_suffix = "/3",
  tooltip_labels_format = "{point.value:.2f}"
)

# Apply height to highcharter object
if (inherits(result, 'highchart')) {
  result <- highcharter::hc_chart(result, height = 550)
}

result
```


:::

