---
title: "{{< iconify ph info >}} About"
format: html
---

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
dashboardr::.page_config(accessibility = TRUE)
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


## About This Demo



This dashboard demonstrates the four loading overlay themes available in dashboardr.

Each page shows a different theme — reload the page to see the overlay animation again.



### Overlay Themes



| Theme | Description |

|-------|-------------|

| `glass` | Semi-transparent frosted effect |

| `light` | Clean white background |

| `dark` | Dark background, ideal for dark themes |

| `accent` | Uses your dashboard's accent color |



### Usage



```r

create_page(

  "My Page",

  overlay = TRUE,

  overlay_theme = "glass",

  overlay_text = "Loading...",

  overlay_duration = 2000

)

```



### Source Code



The full R code that generates this dashboard is available on GitHub:



- [View Source Code](https://github.com/favstats/dashboardr/blob/main/pkgdown/build-overlay-demo.R)



### Learn More



- [dashboardr Documentation](https://favstats.github.io/dashboardr/)

- [Community Gallery](https://favstats.github.io/dashboardr/gallery/)

