---
title: "{{< iconify ph info >}} About"
format: html
---

```{r, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
dashboardr::.page_config(accessibility = TRUE)
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


## About This Demo

This dashboard explores data from the [General Social Survey (GSS)](https://gss.norc.org/),
one of the longest-running surveys in the United States.

Use the sidebar filters on the Explorer page to slice the data by time period
and demographic breakdowns. The dashboard automatically updates charts based on
your filter selections.

### Features Demonstrated

- **Sidebar layout** with radio button filters
- **Stacked bar charts** with percentage and count modes
- **Timeline charts** for tracking trends over time
- **Conditional visibility** (`show_when`) to display different chart types based on filter state
- **Dynamic titles** that update based on selected filters

### Usage

```r
create_page("Explorer", layout = "sidebar") %>%
  begin_sidebar() %>%
    add_input(type = "radio", filter_var = "time_period", ...) %>%
    add_input(type = "radio", filter_var = "breakdown_type", ...) %>%
  end_sidebar() %>%
  add_viz(type = "stackedbar", show_when = ~ time_period != "Over Time") %>%
  add_viz(type = "timeline", show_when = ~ time_period == "Over Time")
```

### Source Code

The full R code that generates this dashboard is available on GitHub:

- [View Source Code](https://github.com/favstats/dashboardr/blob/main/pkgdown/build-sidebar-gss-demo.R)

### Learn More

- [dashboardr Documentation](https://favstats.github.io/dashboardr/)
- [Community Gallery](https://favstats.github.io/dashboardr/gallery/)

