---
title: "{{< iconify ph info >}} About"
format: html
---

```{r, include=FALSE}
# Initialize dashboardr page configuration (CSS/JS for charts)
dashboardr:::.page_config()
```

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| error: false
#| results: 'hide'

# Load dashboardr (includes dplyr, highcharter as dependencies)
library(dashboardr)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)

```


### About This Dashboard


This dashboard demonstrates the **multi-page sidebar** feature in dashboardr.


Each page can have:


- Its own sidebar (or none)


- Different input types (checkbox, radio, select, slider)


- Left or right sidebar position


- Independent filtering for that page's visualizations


- Multi-column grid layouts for checkboxes (using `columns = 2`)


---


### Features Demonstrated


1. **GDP Growth**: Left sidebar with checkbox filters in 2-column grid


2. **Unemployment**: Left sidebar with multi-select dropdown


3. **Combined**: Right sidebar with radio buttons + checkboxes


4. **Slider Demo**: Left sidebar with year slider + checkboxes


---


*Data is simulated for demonstration purposes.*

